### 2022-05-24 Documents/manuscript.tex

\documentclass[twocolumn, 11pt, letterpaper]{article}

% Font
\usepackage[utf8]{inputenc}
\usepackage{MinionPro} 
\input{glyphtounicode}
\pdfgentounicode=1 
\usepackage{microtype}

% Format
\usepackage[letterpaper, margin = 1in]{geometry}
\setlength{\columnsep}{18pt}
\setcounter{secnumdepth}{0}
\usepackage[compact]{titlesec}
\titleformat*{\section}{\centering\normalfont\Large\bfseries}
\titleformat*{\subsection}{\raggedright\normalfont\normalsize\bfseries}
\titleformat*{\subsubsection}{\raggedright\normalfont\normalsize\bfseries\itshape}
\usepackage{authblk}
\renewcommand\Affilfont{\small}
\newcommand{\refsection}{\section{References}}
\providecommand{\tightlist}{\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Links
\usepackage[colorlinks = true, linkcolor = black, urlcolor = black, citecolor = black]{hyperref}

% Figures
\usepackage{graphicx}
\usepackage[labelfont = bf, font = small, labelsep = newline, singlelinecheck = false]{caption}

% Tables
\usepackage{booktabs}
\usepackage{tabularx}

% References
\newenvironment{CSLReferences}[2]{}{}

% Frontmatter
\title{ Double Standards in Judging Collective Action\thanks{} }
\author[1]{Nils K. Reimer}
\affil[1]{University of California, Santa Barbara, United States}
\author[2]{Marija Brankovi\'{c}}
\affil[2]{Singidunum University, Serbia}
\author[3]{Iniobong Essien}
\affil[3]{Leuphana University of L\"{u}neburg, Germany}
\author[4]{Jin X. Goh}
\affil[4]{Colby College, United States}
\author[5]{S\'{e}bastien Goudeau}
\affil[5]{University of Poitiers, France}
\author[6]{N\'{o}ra A. Lantos}
\affil[6]{E\"{o}tv\"{o}s Lor\'{o}nd University, Hungary}
\author[7]{Jenny Veldman}
\affil[7]{KU Leuven, Belgium}

%\date{June 15, 2021}

\begin{document}

\twocolumn[
  \begin{@twocolumnfalse}
    \maketitle
    \begin{abstract}
      \noindent Collective action is a powerful force driving social
change but, at the same time, often sparks contention about what are
acceptable forms of collective action.
    \end{abstract}
  \end{@twocolumnfalse}
  % \renewcommand{\abstractname}{Statement of Relevance}
  % \begin{@twocolumnfalse}
  %   \maketitle
  %   \begin{abstract}
  %   \noindent 
  %   \end{abstract}
  % \end{@twocolumnfalse}
]

{
  \renewcommand{\thefootnote}%
    {\fnsymbol{footnote}}
  \footnotetext[1]{M.B.--J.V. are listed in alphabetical order. This research was supported by a Seedcorn Research Grant from the European Association of Social Psychology.}
}

\noindent Protest movements often spark contention about what actions
are acceptable means of protest. Reactions to Black American athletes
kneeling during the national anthem to protest racist police violence
are a case in point. Kneeling during the national anthem is not violent,
disruptive, or illegal. And yet, only 29\% of White Americans (66\% of
Black Americans) and 11\% of Republicans (59\% of Democrats) considered
it appropriate for Black athletes to protest in this way (YouGov, 2017).
Indeed, Black athletes who had participated in the protests faced a
backlash in public opinion (Lacina, 2020). This and other examples
suggest that \emph{who} the protesters are and \emph{what} they are
protesting influences how acceptable observers judge their actions to
be.

In this paper, we examine double standards in judging collective
action---that is, whether observers judge the same protest actions as
more or less acceptable depending on their own and the protesters' group
memberships (\emph{identity-based double standards}) or on the
protesters' cause and how it aligns with their own ideological positions
(\emph{ideology-based double standards}). In two studies, we first
develop an instrument of 25 controversial protest actions, based in item
response theory, to capture double standards in judging collective
action. In two preregistered experiments, we then use this instrument to
test for double standards in judging protest actions for workers' rights
in the United Kingdom and for or against defunding the police in the
United States. We find evidence for ideology-based, but not for
identity-based, double standards: People judge the same protest actions
to be more acceptable when the protesters' cause aligns with their own
ideological positions---but show no ingroup bias when judging collective
action by ingroup and outgroup members.

\hypertarget{double-standards-in-judging-collective-action}{%
\subsection{Double Standards in Judging Collective
Action}\label{double-standards-in-judging-collective-action}}

Collective action---that is, any action taken by group members to
advance a shared political goal (for similar definitions, see Becker,
2012; van Zomeren, 2016)---is a powerful force driving social change.
For example, Black Lives Matter protests lastingly shifted public
discourse about racial inequity toward antiracist ideas (Dunivin et al.,
2022).

Collective action can take many forms. Psychologists distinguish between
\emph{normative} collective action that seeks to achieve a political
goal while conforming to the norms of the existing social system and
\emph{non-normative} collective action that violates those norms (Wright
et al., 1990). Recent research has shown that this distinction matters
for how people respond to collective action. Shuman et al. (2021) showed
that non-normative, non-violent collective action is more effective than
either normative or violent collective action at gaining concessions
from those most resistant to social change. Feinberg et al. (2020)
demonstrated, however, that observers are less supportive of social
movements that use extreme protest actions. Similarly, Teixeira et al.
(2020) found that advantaged-group members perceive non-normative
collective action by disadvantaged-group members as more damaging to
their ingroup's social image and are thus less supportive of such
action. This means that activists face the dilemma that non-normative
collective action might be most effective at gaining concessions but
that it also risks reducing popular support for a cause. Together, these
studies show that the distinction between normative and non-normative
collective action is psychologically and societally consequential.

Research has, for the most part, relied on ad-hoc distinctions between
what researchers themselves considered normative and non-normative
collective action. In liberal democracies, researchers tend to consider
actions such as signing petitions, voting in elections, or peaceful
protest to be normative and actions such as blocking traffic, damaging
property, or violent protest to be non-normative. But, as the examples
in the first paragraph show, the distinction between normative and
non-normative collective action might instead be in the eye of the
beholder.

In this research, we move this distinction from the researcher's
intuition into the realm of scientific investigation and test for double
standards in where people draw the line between acceptable and
unacceptable forms of collective action. For that purpose, we define a
double standard as judging the same action as more or less acceptable
depending on who is performing the action and for what reason (for an
analogous definition, see Foschi, 2000).\footnote{Feinberg et al. (2020)
  provided incidental evidence for double standards in judging
  collective action as, in a manipulation check, Black participants
  rated Black Lives Matter protests as less extreme than White
  participants and liberal participants rated protest actions for
  progressive causes as less extreme than conservative participants.} In
this section, we derive predictions from social psychological theories
about how identity-based and ideology-based double standards could lead
people to judge the same protest actions as more or less acceptable
depending on who the protesters are and what they are protesting.

\hypertarget{identity-based-double-standards}{%
\subsubsection{Identity-Based Double
Standards}\label{identity-based-double-standards}}

Social Identity Theory (Tajfel \& Turner, 1979; for a recent review, see
Reimer et al., 2022) argues that just as people show self-serving biases
to achieve positive self-esteem, they show ingroup-serving biases to
achieve positive social identities. In other words, they think about
ingroup and outgroup in ``me''--``not me'' terms and favor ingroup
members (``us'') over outgroup members (``not us,'' Brewer, 2007).

One domain in which ingroup bias manifests is judgements about other
people's actions. Hewstone (1990) reviewed research showing that people
make ingroup-serving causal attributions when judging actions by ingroup
and outgroup members, for example, by attributing negative behavior by
outgroup members to internal causes but attributing negative behavior by
ingroup members to external causes. Valdesolo \& DeSteno (2007)
demonstrated that participants judged the same selfish action as less
unfair when performed by themselves or an ingroup member than when
performed by an outgroup member. Other studies (Abrams et al., 2013;
Endevelt et al., 2021; for exceptions, see Mendoza et al., 2014; Pinto
et al., 2010) provided further evidence for ingroup bias in judging
moral transgressions by ingroup and outgroup members.

In the same vain, we propose that ingroup bias results in identity-based
double standards in judging collective action by ingroup and outgroup
members. That is, we hypothesize that people will judge the same protest
actions as more acceptable if the protesters are ingroup members
(\emph{Hypothesis 1a}) or if the cause of the protest aligns with the
ingroup's interests (\emph{Hypothesis 1b}). This hypothesis could
explain why, for example, far more Black than White Americans considered
it appropriate for Black athletes to kneel down to protest police
violence against Black Americans (YouGov, 2017). Notably, this
hypothesis applies equally to historically disadvantaged groups
mobilizing against social injustice and to historically advantaged
groups defending their group's position.

\hypertarget{ideology-based-double-standards}{%
\subsubsection{Ideology-Based Double
Standards}\label{ideology-based-double-standards}}

System Justification Theory (Jost \& Banaji, 1994; for a recent review,
see Jost, 2020) argues that people are motivated to defend, justify, and
bolster the prevailing social, economic, and political system because
doing so serves basic epistemic, existential, and relational needs. Both
advantaged and disadvantaged groups are thought to be motivated to
justify the existing system, although the strength of this motivation
and its expression vary across individuals and situations.

While system justification often leads people to resist social change,
its relationship to collective action depends on the goal of the action.
Jost et al. (2017) argued that research on collective action needs to
distinguish between system-challenging protest aimed at changing an
unequal status quo and system-supporting protest aimed at maintaining or
defending an unequal status quo. Osborne et al. (2019) showed that, for
members of both advantaged and disadvantaged groups, system
justification is associated with supporting system-supporting actions
but with opposing system-challenging actions.

Just as system justification shapes support for, and opposition to,
collective action, we propose that it results in ideology-based double
standards in what actions people consider acceptable forms of collective
action. First, we hypothesize that people will judge the same protest
actions as more acceptable if the protest supports, rather than
challenges, the system (\emph{Hypothesis 2a}). This hypothesis reflects
the assumption that the motivation to justify the system is, to some
extent, universal. Second, we hypothesize that people will judge the
same protest actions as more acceptable if they endorse
system-justifying beliefs and the protest supports the system or if they
reject system-justifying beliefs and the protest challenges the system
(\emph{Hypothesis 2b}). This hypothesis implies an symmetry in how
people with different ideological orientations judge system-supporting
and system-challenging collective action.

As system-justifying beliefs are closely related to political
conservatism (Jost et al., 2003), this hypothesis could explain partisan
differences in what actions liberals and conservatives consider
acceptable means of protest for progressive and conservative causes.
Ideology-based double standards could, for example, explain why far more
Democrats than Republicans considered it appropriate to kneel during the
national anthem to protest racist police violence (YouGov, 2017).

\hypertarget{using-item-response-theory-to-detect-double-standards}{%
\subsection{Using Item Response Theory to Detect Double
Standards}\label{using-item-response-theory-to-detect-double-standards}}

\begin{figure}
\centering
\caption{Item response curves for three hypothetical protest actions}
\includegraphics[scale=1]{../Scale Development/figures/figure-1}
\caption*{\textit{Note.} $\theta_{(z)}$ follows a standard normal distribution and is shown as both $z$-scores (bottom) and percentiles (top).}
\label{fig:f1}
\end{figure}

Item response theory is a a conceptual and statistical framework for
understanding how the characteristics of both items and respondents
shape responses to a set of items (DeMars, 2010). When respondents judge
whether each of a set of actions is an acceptable means of protest (1 =
\emph{yes}, 0 = \emph{no}), a two parameter logistic item response
theory model estimates responses as a function of several latent
(unobserved) parameters:
\[ \Pr ( y_{ij} = 1 ) = \text{logit}^{-1} ( \alpha_i ( \theta_j + \beta_i ) ) \]
where \(\Pr ( y_{ij} = 1 )\) is the probability that respondent \(j\)
considers action \(i\) an acceptable means of protest, \(\theta_j\)
estimates how \emph{accepting} respondent \(j\) is of various protest
actions, \(\beta_i\) estimates how \emph{acceptable} action \(i\) is,
and \(\alpha_i\) estimates how \emph{discriminating} action \(i\) is.

Figure \ref{fig:f1} illustrates those relationships for three
hypothetical protest actions. \(\theta_{(z)}\) is the
\emph{z}-standardized tendency for a participant to be more accepting of
various actions so that \(\beta_i\) determines the probability that the
average respondent (\(\theta_{(z)} = 0\)) considers an action an
acceptable means of protest. Action 2 (\(\beta_2 = 1.28\)) is more
acceptable than Actions 1 and 2 (\(\{ \beta_2, \beta_3 \} = 0\)) as the
average respondent is more likely to consider the former (\(\Pr = .78\))
than the latter (\(\Pr = .50\)) acceptable means of protest. Action 3
(\(\alpha_3 = 3.60\)) is more discriminating than Actions 1 and 2
(\(\{ \alpha_1, \alpha_2 \} = 1\)) as the slope for the relationship
between how accepting a respondent is and how likely they are to
consider an action acceptable is steeper for the former action. This
means that knowing whether someone considers Action 3 acceptable
provides more information about them than knowing whether they consider
Actions 1 or 2 acceptable.

In this research, item response theory serves two purposes. First, item
response theory provides a formal and statistical definition of double
standards in judging collective action as respondents being more
accepting of the same protest actions depending on who the protesters
and what they are protesting. In Experiments 1 and 2, we operationalize
double standards as the differences in \(\theta_{(z)}\) between
conditions which, as \(\theta_{(z)}\) is \(z\)-standardized, correspond
to Cohen's \(d\) effect sizes. Second, item response theory provides a
statistical framework for scale development. In Studies 1 and 2, we
select the most discriminating and informative actions to build an
instrument for capturing double standards in the experiments.

\hypertarget{purpose-of-the-present-research}{%
\subsection{Purpose of the Present
Research}\label{purpose-of-the-present-research}}

Our research serves two purposes. The first purpose is to develop an
instrument to measure where people draw the line between acceptable and
unacceptable forms of collective action. In Study 1, we ask participants
to generate more and less extreme protest actions and compile a pool of
protest actions for scale development. In Study 2, we ask participants
to rate how acceptable they consider each of those protest action to be
and use item response theory to select the most discriminating and
informative protest actions for our instrument. In so doing, we build an
instrument that enables both present and future research to capture
double standards in judging collective action. In this way, our research
situates the distinction between normative and non-normative collective
action in the eye of the beholders and opens it to scientific
investigation.

The second purpose of our research is to test for double standards in
judging collective action---that is, whether people judge the same
protest actions as more or less acceptable depending on their own and
the protesters' group memberships (Hypothesis 1a, 1b) or on the
protesters' cause and how it aligns with their own ideological positions
(Hypotheses 2a, 2b). In Experiment 1, we use the instrument developed in
Studies 1 and 2 to test for double standards in how acceptable
participants consider collective action for workers' rights in the
United Kingdom. By varying both the participants' and protesters' group
memberships (working/middle class), Experiment 1 compares the relative
evidence for Hypotheses 1a and 2a. Extending Experiment 1, Experiment 2
considers both system-challenging and system-defending collective action
in the context of the 2020 Black Lives Matter protests in the United
States. By varying the protest's cause (for/against defunding the
police) as well as the participants' and protesters' group memberships
(Black/White), Experiment 2 compares the relative evidence for
Hypotheses 1a, 1b, 2a, and 2b. Together, the preregistered experiments
provide a comprehensive test of the hypothesized identity- and
ideology-based double standards in judging collective action.

\hypertarget{scale-development}{%
\section{Scale Development}\label{scale-development}}

\begin{figure*}[!t]
\centering
\caption{Examples of protest actions rated in Study 2}
\includegraphics[scale=1]{../Scale Development/figures/figure-2}
\caption*{\textit{Note.} Percentages reflect the proportion of participants who thought that an action was \textit{sometimes}, \textit{often}, or \textit{always} an acceptable means to a advance a cause. Darker bars mark actions that were included in the final scale.}
\label{fig:f2}
\end{figure*}

\hypertarget{study-1}{%
\subsection{Study 1}\label{study-1}}

In Study 1, we compiled protest actions from participants' responses and
other sources. We recruited 60 participants from the Prolific subject
pool, all of whom were citizens of the UK or the US. To increase the
socioeconomic diversity of our sample, we recruited 30 non-students
without a university degree, 15 non-students with a university degree,
and 15 current university students.

Participants first read an accessible definition of what a social group
is and that collective action is any action group members take to
promote the interests of their social group. Participants were then
asked to name between five and ten actions that fit that definition.
Participants were encouraged to think of actions that varied in how
acceptable they were in their opinion. We recoded responses into a
smaller set of unique actions, which we supplemented with protest
actions from the psychological and political science literature (e.g.,
Sharp, 1973). This process resulted in 72 actions that varied in how
acceptable we would expect them to be. For details, see Supplemental
Online Material (SOM).

\hypertarget{study-2}{%
\subsection{Study 2}\label{study-2}}

In Study 2, we measured how acceptable participants judged the actions
from Study 1 to be and applied item response theory to develop an
instrument to capture double standards in judging collective action. We
recruited 158 participants (\(\textit{Mdn} = 30\) years, age range:
18--68 years; 103 women, 52 men, 2 other, 1 prefer not to say) from the
Prolific subject pool, all of whom were citizens of the UK or the US. To
increase the socioeconomic diversity of our sample, we recruited 80
non-students without a university degree, 37 non-students with a
university degree, and 41 current university students. We excluded 15
participants who failed an attention check, leaving a final sample of
143 participants for our analyses.

Participants again read an accessible definition of collective action.
Participants were then asked to think of different causes and
circumstances and to rate how often a given action would be an
acceptable means for a group to advance one of these causes (1 =
\emph{never}, 2 = \emph{rarely}, 3 = \emph{sometimes}, 4 = \emph{often},
5 = \emph{always}). In addition, participants rated how disruptive,
violent, and extreme they considered an action to be (1 = \emph{not at
all}, 4 = \emph{very}) and how positive or negative they felt, in
general, about an action (1 = \emph{very positive}, 5 = \emph{very
negative}). Each participant rated 20 of the 72 actions from Study 1 so
that each action was rated by 29--53 participants.

We estimated a graded response model (Bürkner, 2021; Samejima,
1997)---an item response theory model for ordinal response
variables---for participants' ratings of how often an action would be an
acceptable form of collective action.\footnote{Ratings of how often a
  given action would be an acceptable form of collective action were
  strongly and negatively correlated with how disruptive, violent, and
  extreme participants considered the action to be and how negative they
  felt about it (Table S3).} Based on the results, we selected the 25
most discriminating, informative, and relevant protest actions to form
an instrument for measuring double standards in judging collective
action in Experiments 1 and 2. We selected controversial, and thus
diagnostic, actions (e.g., disrupting traffic) over actions that almost
all respondents considered acceptable (e.g., signing petitions) or
unacceptable (e.g., political violence; Figure \ref{fig:f2}). For
detailed ratings and results, see SOM.

\hypertarget{experiment-1}{%
\section{Experiment 1}\label{experiment-1}}

\ldots

\hypertarget{method}{%
\subsection{Method}\label{method}}

We preregistered the sample size as well as all hypotheses,
inclusion/exclusion criteria, measures, and manipulations.\footnote{\url{https://osf.io/24wrx/?view_only=8560eb60286149498f8da48bc9f9d99b}}
We made all materials, data, and analysis scripts available
online.\footnote{\url{https://osf.io/d3yev/?view_only=40782034017c40f0bcecb1cc87760b62}}
We followed sample size recommendations for item response theory models
(DeMars, 2010, p. 36), planning to recruit 500 participants.

\hypertarget{study-design}{%
\subsubsection{Study Design}\label{study-design}}

We used a 2 (quasi-experimental: higher-status/lower-status
participants) \(\times\) 2 (experimental: higher-status/lower-status
protesters) between-subjects design to test our hypotheses.

\hypertarget{participants}{%
\subsubsection{Participants}\label{participants}}

We recruited 515 participants from the Prolific subject pool who were UK
citizens, 25 years old or older, and not current students.\footnote{We
  had preregistered a sample of 500 participants, before exclusions, but
  included another 15 participants who completed the study without
  returning an approval code to the Prolific platform. Participants
  received, on average, \pounds 9.00 (\$10.93) per hour of
  participation.} As preregistered, we excluded 71 participants who
failed an attention check. This resulted in a final sample of 443
participants (\(\textit{Mdn} = 41\) years, age range: 25--76 years; 272
women, 171 men, 1 non-binary). Of these, 210 participants considered
their past, current, or future jobs to be working-class jobs.
Participants in this lower-status group did not have a university degree
and placed themselves on the bottom three ranks of the subjective
socioeconomic status ladder. Another 233 participants considered their
past, current job, or future jobs to be middle-class/professional jobs.
Participants in this higher-status group had at least an undergraduate
degree and placed themselves on the top four ranks of the subjective
socioeconomic status ladder.

\hypertarget{procedure}{%
\subsubsection{Procedure}\label{procedure}}

We used a screening survey to recruit participants who satisfied our
preregistered inclusion criteria for the lower-status and higher-status
groups. For the lower-status group, we recruited 475 participants who
did not have a university degree and placed themselves on the bottom
three ranks of a subjective socioeconomic status ladder. For the
higher-status group, we recruited 400 participants who had at least an
undergraduate degree and placed themselves on the top four ranks of the
subjective socioeconomic status ladder.

In the screening survey, participants read an accessible definition of
what working-class and middle-class/professional jobs are (for details,
see SOM). Participants then answered, among other questions, whether
they considered their current job---or the jobs they had had in the past
or expected to have in the future---to be a working-class job or a
middle-class/professional job. As preregistered, we excluded
participants from the lower-status group who did not respond
``working-class job'' and participants from the higher-status group who
did not respond ``middle-class/professional job''.

We recruited 500 participants from the remaining 687 participants, 250
from the lower-status group and 250 from the higher-status group.
Participants were randomly assigned to read a vignette about a
government bill affecting either people in working-class jobs
(lower-status protesters) or people in professional jobs (higher-status
protesters). Participants in both conditions were instructed to
carefully read the vignette and to try to imagine what it would be like
if this situation was real. Participants in the lower-status protestprs
condition read the following introduction:

\begin{quote}
The government, though not necessarily the current government, is going
to introduce a bill that will mostly affect people in working-class
jobs. Working-class jobs, in this case, are jobs done by skilled,
semi-skilled, unskilled manual workers or by casual workers. These are
jobs that do not usually require a university degree. Other jobs are
unlikely to be affected.
\end{quote}

\noindent Participants in the higher-status protesters condition instead
read the following introduction:

\begin{quote}
The government, though not necessarily the current government, is going
to introduce a bill that will mostly affect people in professional jobs.
Professional jobs are administrative, managerial, or other jobs that
usually require a university degree. Other jobs are unlikely to be
affected.
\end{quote}

\noindent Participants in both conditions then read an almost
identically worded paragraph:

\begin{quote}
This government measure would make it easier for companies to hire
workers during economic growth and to lay off workers during an economic
crisis. As a consequence, companies would be able to fire employees with
little notice and without giving a reason. Trade unions are opposed to
the measure. They argue that the bill would compromise job security, and
prevent employees from challenging harassment or other abuse without the
fear of being fired. People in {[}working-class/professional{]} jobs are
particularly at risk, and there is a rise in tension and outrage among
them.
\end{quote}

On the next pages, participants completed all remaining measures. On the
final page, we asked participants to recall who the people most affected
by the fictitious government bill were. As preregistered, we excluded
all participants whose response did not qualitatively match their
experimental condition.

\hypertarget{measures}{%
\subsubsection{Measures}\label{measures}}

We measured the outcome variable by asking participants to decide, for
each of 25 protest actions presented in a randomized order, whether they
thought this action was ``an acceptable means for people in
{[}working-class/professional{]} jobs to protest against the government
bill'' (1 = \emph{yes}, 0 = \emph{no}; see Table A1).

We assessed reactions to the vignette by asking participants how
outraged they would be if the government were to introduce this bill in
real life, to what extent it would affect them personally, and to what
extent the it would affect people like them (1 = \emph{not at all}, 7 =
\emph{very much}). We also asked participants to what extent they
identified with people with the kinds of jobs most affected by the
proposed bill (1 = \emph{not at all}, 7 = \emph{very much}) and how
often, if at all, they had participated in protest actions such as the
ones we asked about (1 = \emph{never}, 5 = \emph{very often}).

We measured social dominance orientation with the eight-item
\(\text{SDO}_{7(s)}\) scale (Ho et al., 2015), for example, ``it is
unjust to try to make groups equal'' (1 = \emph{strongly oppose}, 7 =
\emph{strongly favour}; McDonald's \(\omega = .89\)). A confirmatory
factor analysis model in which all items loaded onto a single factor
showed acceptable fit,
\(\chi^2 (20) = 108.69; \text{CFI} = 0.92; \text{TLI} = 0.88; \text{RMSEA} = 0.11, [0.09, 0.14]\).

We measured system-justifying beliefs with eight items (adapted from Kay
\& Jost, 2003), for example, ``in general, I find society to be fair''
(1 = \emph{strongly disagree}, 7 = \emph{strongly agree}; McDonald's
\(\omega = .89\)). A confirmatory factor analysis model in which all
items loaded onto a single factor showed acceptable fit,
\(\chi^2 (20) = 77.37; \text{CFI} = 0.94; \text{TLI} = 0.92; \text{RMSEA} = 0.09, [0.07, 0.12]\).

In the screening survey, we recorded participants' gender, age,
nationality, student status, and employment status. We also included two
three-item scales measuring social identification with people in
working-class and middle-class/professional jobs (adapted from Becker et
al., 2011) and a one-item semantic differential scale measuring
political orientation (``People often describe their political
orientation as left-wing or right-wing. On a scale from left to right,
where would you position yourself?''; 1 = \emph{left}, 7 =
\emph{right}).

\hypertarget{results}{%
\subsection{Results}\label{results}}

\begin{figure*}[!t]
\centering
\caption{Results from the preregistered (\textbf{A}) and non-preregistered (\textbf{B}) analyses for Experiment 1}
\includegraphics[scale=1]{../Experiment 1/figures/figure-3}
\caption*{\textit{Note.} HS = Higher Status, LS = Lower Status. $\theta_{(z)}$ is the \textit{z}-standardized tendency to consider more controversial actions acceptable means of protest. (\textbf{A}) $p$ and $n$ are, respectively, the predicted proportion and number of actions a participant would consider acceptable means of protest in each condition. Bars enclose the 95\% most plausible estimates. (\textbf{B}) Ribbons enclose, from the darkest to the lightest shade, the 50\%, 80\%, 95\%, and 99\% most plausible estimates.}
\label{fig:f3}
\end{figure*}

\hypertarget{reactions-to-the-manipulation}{%
\subsubsection{Reactions to the
Manipulation}\label{reactions-to-the-manipulation}}

Participants with professional jobs thought that they would be more
outraged if the government were to introduce a bill affecting people
with similar jobs (\(M = 6.00, SD = 1.05\)) than by a bill affecting
people with working-class jobs (\(M = 5.38, SD = 1.28\); Cohen's
\(d = 0.50\)). Conversely, participants with working-class jobs thought
that they would be more outraged by a bill affecting people in similar
jobs (\(M = 6.26, SD = 1.05\)) than by a bill affecting people in
professional jobs (\(M = 5.61, SD = 1.47\); \(d = 0.52\)). Participants
thought that a bill affecting people with similar jobs would affect them
personally (\(d = 1.14\)) and people like them (\(d = 1.33\)) more than
a bill affecting people with other jobs. Participants identified to a
greater extent with people in similar jobs (\(M = 6.21, SD = 1.12\))
than with people in other jobs (\(M = 4.06, SD = 1.65\); \(d = 1.22\)).
Overall, participants' reactions suggested that the experimental design
worked as intended and that participants in the lower-status and
higher-status groups understood the vignette in ingroup--outgroup terms.

\hypertarget{preregistered-analyses}{%
\subsubsection{Preregistered Analyses}\label{preregistered-analyses}}

To test our hypotheses, we estimated a two-parameter logistic item
response theory model (Bürkner, 2021) with participants' responses to
the question whether they thought each action was an acceptable means of
protest as the outcome variable. For each action \(i\), the model
estimated how acceptable (\(\beta_i\)) and discriminating (\(\alpha_i\))
that protest action was. For each participant \(j\), the model estimated
their unique propensity to consider protest actions acceptable
(\(\theta_j\)). In addition, the model estimated how accepting
participants were of controversial protest actions as a function of
three dummy-coded variables that encoded the group status of the
protesters and the participant.

We estimated this model in CmdStan (Gabry \& Cesnovar, 2021; Stan
Development Team, 2021) using Bayesian statistical methods. Bayesian
inference involves choosing a likelihood function and prior
distributions. A likelihood function links the observed data to the
model parameters and states how likely the observed data are given
different values of said model parameters. Prior distributions state how
plausible different values of said model parameters are before
considering the observed data. Bayesian inference applies Bayes' theorem
to update prior distributions in light of the observed data to produce
posterior distributions. In contrast to \(p\)-values and confidence
intervals, posterior distributions have a straight-forward
interpretation as stating how plausible different values of the model
parameters are given the observed data.

Our model derived the likelihood of participants' responses from a
Bernoulli likelihood function with a logistic regression equation
linking the two item parameters (\(\alpha_i\), \(\beta_i\)), the one
participant parameter (\(\theta_j\)), and the three regression
coefficients to the observed data. To identify the model, we constrained
\(\theta_j\) to have a mean of zero and constrained \(\alpha_i\) to have
a fixed mean and to be non-negative. We used partial pooling to estimate
\(\alpha_i\), \(\beta_i\), and \(\theta_j\). Our model assigned weakly
informative prior distributions (Gelman et al., 2017) to all model
parameters.\footnote{Our model assigned \(\text{Half-Cauchy} (0, 3)\)
  prior distributions to the standard deviations of \(\alpha_i\),
  \(\beta_i\), and \(\theta_j\) and \(\text{Normal} (0, 3)\) prior
  distributions to all other model parameters.} We report point
estimates, based on the median of posterior samples, and uncertainty
intervals, based on the quantiles of posterior samples, that enclose the
95\% most plausible estimates.

Figure \ref{fig:f3} (A) shows estimates for each combination of the
protesters' and the participants' group membership. Overall, we found
that participants' responses depended on both the protesters' and the
participants' group status---but not in the directions predicted by our
hypotheses. Contradicting Hypothesis 1, participants did not consider
protest actions performed by their ingroup to be more acceptable, on
average, than the same actions performed by the relevant outgroup
(Cohen's \(d = 0.00, [-0.20, 0.20]\)). Instead, we found that
participants from higher-status backgrounds considered protest actions
performed by both lower-status (\(d = 0.34, [0.07, 0.59\){]}) and
higher-status (\(d = 0.33, [0.03, 0.63]\)) protesters to be, on average,
more acceptable than participants from lower-status backgrounds.
Contradicting Hypothesis 2, participants considered protest actions
performed by higher-status protesters to be less acceptable, on average,
than the same actions performed by lower-status protesters
(\(d = -0.20, [-0.40, -0.00]\)).

\hypertarget{non-preregistered-analyses}{%
\subsubsection{Non-Preregistered
Analyses}\label{non-preregistered-analyses}}

We explored to what extent participants' ideological orientations
influenced how acceptable they considered various protest actions to be.
To that end, we estimated another two-parameter logistic item response
model that estimated participants' responses as a function of the
participant's and the protesters' group status and of the participants'
political orientation, social dominance orientation, and
system-justifying beliefs. We used factor scores to quantify social
dominance orientation and system-justifying beliefs and standardized all
new predictor variables. We report standardized regression coefficients
that quantify by how many standard deviations a participant's propensity
to consider more collective actions acceptable increases for each
additional standard deviation of the predictor variable.

Figure \ref{fig:f3} (B) the \(z\)-standardized propensity to consider
more controversial protest actions acceptable as a function of the three
ideological orientation variables. We found that participants who
reported a more right-wing political orientation
(\(\beta_{xy} = -0.25, [-0.37, -0.15]\)) and who expressed more
agreement with system-justifying beliefs
(\(\beta_{xy} = -0.20, [-0.30, -0.10]\)) tended to find fewer collective
actions to be acceptable. In contrast, we found that, after controlling
for the other two variables, social dominance orientation was not
associated with participants' judgements about how acceptable various
collective actions are (\(\beta_{xy} = -0.05, [-0.15, 0.05]\)). Overall,
these findings suggest that people who are right-wing and endorse
system-justifying beliefs tend to find various collective actions to be
less acceptable than people who are left-wing and do not endorse
system-justifying beliefs.\footnote{We ran additional analyses to
  explore how these associations differed across experimental
  conditions. We found uncertainty intervals for these associations to
  overlap across conditions, suggesting that we did not have enough data
  to differentiate these varying effects.}

\hypertarget{discussion}{%
\subsection{Discussion}\label{discussion}}

\hypertarget{experiment-2}{%
\section{Experiment 2}\label{experiment-2}}

\begin{itemize}
\tightlist
\item
  Explain why defunding the police (public support for BLM, so needed to
  be more controversial)
\end{itemize}

In Experiment 2, we tested five hypotheses. We tested two predictions
derived from social identity theory. First, we hypothesize that
observers are more likely to judge an action as an acceptable means of
protest when the protesters are ingroup members than when they are
outgroup members (Hypothesis 1a). Second, we hypothesize that observers
are more likely to judge an action as an acceptable means of protest
when the goal of the protests would benefit the observers' ingroup
(Hypothesis 1b). We will test two predictions derived from system
justification theory. First, we hypothesize that observers are more
likely to judge an action as an acceptable means of protest when the
goal of the protest is to support rather than challenge the system
(Hypothesis 2a). Second, we hypothesize that observers who endorse
system-justifying beliefs are more likely to consider system-supporting
actions acceptable means of protest-and less likely to consider
system-challenging actions acceptable means of protest-than observers
who do not endorse system-justifying beliefs (Hypothesis 2b). In
addition, we will test an alternative hypothesis. We hypothesize that
observers are more likely to consider actions as acceptable means of
protest when they support the cause of the protesters than when they
oppose their cause (Hypothesis 3). We note that Hypotheses 1a, 1b, and
2a make competing predictions about differences between conditions while
Hypotheses 2b and 3 need not contradict each other or any of the other
hypotheses.

\hypertarget{method-1}{%
\subsection{Method}\label{method-1}}

We preregistered the sample size as well as all hypotheses,
inclusion/exclusion criteria, measures, and manipulations.\footnote{\url{https://osf.io/skxjt/?view_only=d8ce44a700884b5ab36b64ef08f833a1}}
We made all materials, data, and analysis scripts available
online.\footnote{\url{https://osf.io/d3yev/?view_only=40782034017c40f0bcecb1cc87760b62}}
As reported in the preregistration, we ran simulations, using data from
Experiment 1, to determine that a sample size of \(N = 1,600\)
(\(n = 200\) per condition) was sufficient to detect even small
differences between conditions (\(0.09 < \text{Cohen's}~d < 0.16\)).

\hypertarget{study-design-1}{%
\subsubsection{Study Design}\label{study-design-1}}

We used a 2 (quasi-experimental: Black/White participants) \(\times\) 2
(experimental: Black/White protesters) \(\times\) 2 (experimental:
for/against defunding the police) between-subjects design to test our
hypotheses.

\hypertarget{participants-1}{%
\subsubsection{Participants}\label{participants-1}}

We recruited 1,773 Black and White American participants from the
Prolific subject pool who were 18 years old or older, lived in the US,
and were US citizens.\footnote{Data were collected between January 15
  and 28, 2021. Participants received, on average, \$12.89 per hour of
  participation.} As preregistered, we excluded 173 participants who
failed at least one of three attention checks or who reported a
different ethnic background than they had reported in the Prolific
prescreening questionnaire. This resulted in a final sample of 1,600
participants (\(\textit{Mdn} = 31\) years, age range: 18--84 years; 864
women, 708 men, 27 sex/gender diverse) of whom 800 identified as Black
and 800 identified as White.

\hypertarget{procedure-1}{%
\subsubsection{Procedure}\label{procedure-1}}

Participants read the following paragraphs:

\begin{quote}
In 2020, police officers killed Breonna Taylor and George Floyd. Many
were outraged that police officers had, once again, killed unarmed Black
Americans. Across the United States, people called for changes to
prevent future police violence.
\end{quote}

\begin{quote}
Some argue that, to end police violence, we should take money away from
police departments. Reducing police funding would mean fewer police
officers on the street. Fewer police officers would mean fewer
opportunities for them to turn violent. Reducing police funding would
also leave more money for other services. Proponents argue for
reallocating police funding to social services, housing, and education.
Doing so would keep communities safer with fewer police officers. We
refer to this position as ``defunding the police''. This position
differs from ``reforming the police'' which might mean increasing police
funding and also differs from ``abolishing the police'' which means
disbanding police departments altogether.
\end{quote}

\noindent Participants were then asked to answer whether the text had
been about ``reforming the police'', ``defunding the police'', or
``abolishing the police''. If they selected the wrong answer, they were
instructed to reread the text and select the right answer. On the next
page, participants stated whether they supported or opposed the proposed
solution.

Participants in the system-challenging protest condition then read a
text about protesters in support of defunding the police:

\begin{quote}
Earlier, you read about defunding the police as a possible solution to
end police violence. Some local residents want to protest for defunding
the police. They argue that reducing police funding would prevent police
violence.
\end{quote}

\noindent Participants in the system-supporting protest condition
instead read about protesters rallying against defunding the police:

\begin{quote}
Some local residents want to protest against defunding the police. They
argue that reducing police funding would mean fewer police officers
serving their community.
\end{quote}

\noindent Participants then read either that ``most of the protesters
are Black'' or that ``most of the protesters are White''. Participants
again had to correctly answer multiple choice questions about the text
(``Are the protesters for or against defunding the police?'', ``Who are
the protesters?'') before being able to proceed.

On the next pages, participants completed all remaining measures. On the
final page, participants responded to three attention checks: ``In this
study, you first read about a proposed solution to police violence. What
was it?'' (\emph{reforming the police}, \emph{defunding the police},
\emph{abolishing the police}); ``In this study, you then read about
protesters. Were these protesters for or against the proposed
solution?'' (\emph{for}, \emph{against}); and ``Were most of the
protesters Black or White?'' (\emph{Black}, \emph{White}). As
preregistered, we excluded participants who gave an answer inconsistent
with their assigned experimental condition.

\hypertarget{measures-1}{%
\subsubsection{Measures}\label{measures-1}}

We measured the outcome variable by asking participants to think about
the protesters they had read about and to decide, for each of 25 protest
actions presented in a randomized order, whether they thought this
action was ``an acceptable means for them to protest {[}for/against{]}
defunding the police'' (1 = \emph{yes}, 0 = \emph{no}). We replaced some
actions from Experiment 1 because they either did not fit the study
context or had been considered acceptable by almost all participants
(see Table B1).

We measured system-justifying beliefs with eight items (adapted from Kay
\& Jost, 2003), for example, ``in general, I find society to be fair''
(1 = \emph{strongly disagree}, 7 = \emph{strongly agree}; McDonald's
\(\omega = .91\)). A confirmatory factor analysis model in which all
items loaded onto a single factor showed acceptable fit,
\(\chi^2 (28) = 4309.15; \text{CFI} = 0.95; \text{TLI} = 0.93; \text{RMSEA} = 0.11, [0.10, 0.12]\).

We measured support for defunding the police with one item: ``Do you
support or oppose defunding the police?'' (1 = \emph{strongly oppose}, 5
= \emph{strongly support}).

We included additional measures to describe the sample, to describe
reactions to the manipulation, or to use in non-preregistered analyses.
In addition to demographic questions, we asked participants how outraged
they were about recent incidents of police violence against Black
Americans, to what extent they identifed with the protesters described
in the study, and to what extent they identifed with their racial
ingroup (1 = \emph{not at all}, 7 = \emph{very much}). We also asked how
often, if at all, participants had participated in protest actions such
as the ones we had asked about (1 = \emph{never}, 5 = \emph{very often})
and whether they has participated in protests for reforming, defunding
or abolishing the police; against reforming, defunding or abolishing the
police; or in neither. We measured political orientation with a one-item
semantic differential scale: ``People often describe their political
orientation as liberal or conservative. On a scale from liberal to
conservative, where would you position yourself?''; 1 = \emph{liberal},
7 = \emph{conservative}).

\hypertarget{results-1}{%
\subsection{Results}\label{results-1}}

\begin{figure*}[!t]
\centering
\caption{Results from the preregistered analyses for Experiment 2}
\includegraphics[scale=1]{../Experiment 2/figures/figure-4}
\caption*{\textit{Note.} Against = Protesters oppose defunding the police. For = Protesters support defunding the police. $\theta_{(z)}$ is the \textit{z}-standardized tendency to consider more controversial actions acceptable means of protest. Ribbons enclose, from the darkest to the lightest shade, the 50\%, 80\%, 95\%, and 99\% most plausible estimates.}
\label{fig:f4}
\end{figure*}

\begin{figure*}[!t]
\centering
\caption{Predictions from the preregistered (system justification) and non-preregistered (political orientation) analyses for Experiment 2}
\includegraphics[scale=1]{../Experiment 2/figures/figure-5}
\caption*{\textit{Note.} Against = Protesters oppose defunding the police. For = Protesters support defunding the police. $\theta_{(z)}$ is the \textit{z}-standardized tendency to consider more controversial actions acceptable means of protest.}
\label{fig:f5}
\end{figure*}

\hypertarget{reactions-to-the-manipulation-1}{%
\subsubsection{Reactions to the
Manipulation}\label{reactions-to-the-manipulation-1}}

As expected, Black participants (\(M = 6.18, \textit{SD} = 1.31\))
reported being more outraged about recent incidents of police violence
than White participants (\(M = 5.56, \textit{SD} = 1.61\); Cohen's
\(d = 0.42\)). Black participants strongly identified with their ingroup
(\(M = 6.46, \textit{SD} = 1.09\)) and, across experimental conditions,
identified more with Black protesters (\(M = 4.66, \textit{SD} = 1.99\))
than with White protesters (\(M = 4.03, \textit{SD} = 2.16\);
\(d = 0.30\)). In contrast, White participants identified less strongly
with their ingroup (\(M = 4.85, \textit{SD} = 1.53\); \(d = 1.21\)) and
did not identify more with White protesters
(\(M = 3.58, \textit{SD} = 2.01\)) than with Black protesters
(\(M = 3.81, \textit{SD} = 1.77\); \(d = -0.12\)). Both White
(\(d = 0.67\)) and Black (\(d = 0.53\)) participants tended to identify
more with protesters protesting \emph{for} defunding the police. On
average, participants tended to describe their political orientation as
moderately liberal (\(M = 2.92, \textit{SD} = 1.73\)) but, as expected,
were divided in their support for defunding the police
(\(M = 3.37, \textit{SD} = 1.42\)).

\hypertarget{preregistered-analyses-1}{%
\subsubsection{Preregistered Analyses}\label{preregistered-analyses-1}}

To test our hypotheses, we estimated a series of two-parameter logistic
item response theory models with participants' responses to the question
whether they thought each action was an acceptable means of protest as
the outcome variable. Our models differed from the models used in
Experiment 1 in two ways. First, we estimated the item parameters
(\(\alpha_i\), \(\beta_i\)) as \emph{correlated} varying effects.
Second, we used varying, rather than fixed, effects to estimate
differences between the eight conditions. This resulted in partial
pooling---condition-wise estimates were shrunk towards each other---and
allowed us to compare conditions without multiple comparison problems
(Gelman et al., 2012). Our models assigned weakly informative prior
distributions to all model parameters.\footnote{Our model assigned
  \(\text{LKJ} (2)\) prior distributions to the Cholesky-transformed
  correlation matrices for varying effects and
  \((\text{Half-})\text{Cauchy} (0, 5)\) prior distributions to all
  other model parameters.} Figure \ref{fig:f4} shows results from three
preregistered models we estimated to test our hypotheses.

Model 1 (M1) estimated varying intercepts for the eight conditions to
test whether, in line with Hypotheses 1a, 1b, or 2a, participants'
responses depended on their own group membership, the protesters' group
membership, or the protesters' cause. Contradicting Hypothesis 1a,
participants did not consider protest actions performed by their ingroup
to be more acceptable, on average, than the same actions performed by
the relevant outgroup (\(d = 0.01, [-0.04, 0.06]\)). Contradicting
Hypothesis 1b, participants did not consider protest actions for a cause
that was nominally aligned with their ingroup's interests to be more
acceptable, on average, than protest actions for a cause not aligned
with their ingroup's interests (\(d = -0.01, [-0.06, 0.04]\)).
Contradicting Hypothesis 2a, participants did not consider protest
actions for a system-defending cause to be more acceptable, on average,
than protest actions for a system-challenging cause
(\(d = -0.14, [-0.20, -0.09]\)). Instead, we found that, on average,
White participants considered all protest actions to be more acceptable
than Black participants (\(d = 0.09, [0.04, 0.14]\)) and both Black
(\(d = 0.07, [0.03, 0.10]\)) and White (\(d = 0.08, [0.04, 0.11]\))
participants considered the same actions to be more acceptable when
protests were for, rather than against, defunding the police. As in
Experiment 1, we thus found that participants' responses depended on
both the participants' and the protesters' group memberships---but not
in the directions predicted by our hypotheses.

Model 2 (M2) extended Model 1 by estimating participants' responses as a
function of their \emph{z}-standardized endorsement of system-justifying
beliefs. As preregistered, we modeled this relationship with two fixed
effects, one estimating the effect of system-justifying beliefs on
judgements about system-defending protest actions and one estimating
their effect on judgements about system-challenging protest actions, and
one varying effect estimating its variance across conditions. Supporting
Hypothesis 2b, participants who \emph{rejected} system-justifying
beliefs were \emph{more} likely to consider system-challenging protest
actions (for defunding the police) to be acceptable means of protest
(\(\beta_{xy} = 0.38, [0.16, 0.58]\)). We did not, however, find
evidence for the ideological symmetry implied by Hypothesis 2b since
participants who \emph{endorsed} system-justifying beliefs were
\emph{not} more likely to consider system-defending protest actions
(against defunding the police) to be acceptable means of protest
(\(\beta_{xy} = -0.18, [-0.40, 0.02]\)).

Figure \ref{fig:f5} shows the predicted pattern of condition-wise
differences underlying those fixed effects. Participants who endorsed
system-justifying beliefs tended to consider system-challenging and
system-defending protest actions to be equally unacceptable. In
contrast, participants who rejected system-justifying beliefs evinced
ideology-based double standards: In line with Hypothesis 2b, they
considered the same protest actions to be more acceptable if the
protesters challenged the system or, to a lesser extent, if the
protesters were from the disadvantaged group.

Model 3 (M3) extended Model 1 by estimating participants' responses as a
function of their self-reported support for the cause of the protest. We
recoded participants' responses to create a predictor variable that
encoded support for defunding the police when protesters supported
defunding the police and opposition to defunding the police when
protesters opposed defunding the police. As preregistered, we modeled
this relationship as a monotonic effect (Bürkner \& Charpentier, 2020)
that estimated the average change in the outcome variable across
predictor categories as well as how much of this change occurred between
each of the four pairs of adjacent predictor categories.\footnote{Our
  model assigned a Dirichlet prior, \(\alpha = {1, 1, 1, 1}\), to the
  proportions of the overall change that was expected to occur between
  each of the four pairs of predictor categories.} Contradicting
Hypothesis 3, participants who supported the protesters' cause did not,
on average, consider the same protest actions to be more acceptable than
participants who opposed the protesters' cause
(\(\beta_{y} = 0.14, [-0.12, 0.40]\)). Ergo, our results did not support
the alternative hypothesis that ideology-based double standards can be
reduced to support for, or opposition to, the cause of a protest.

\hypertarget{non-preregistered-analyses-1}{%
\subsubsection{Non-Preregistered
Analyses}\label{non-preregistered-analyses-1}}

We reran Model 3 with political orientation, instead of system
justification, as an alternative predictor variable. Our results
mirrored results from the preregistered analyses: More liberal
participants were more likely to consider protest actions for defunding
the police to be acceptable (\(\beta_{xy} = 0.41, [0.17, 0.62]\)) but
more conservative participants were not more likely to consider protest
actions against defunding the police to be acceptable
(\(\beta_{xy} = -0.16, [-0.41, 0.04]\)). As Figure 5 shows, conservative
participants tended to consider all protest actions to be equally
unacceptable while liberal participants considered the same protest
actions to be more acceptable if protesters rallied around a progressive
cause. Our non-preregistered analyses thus found evidence for
ideology-based double standards in judging collective action with a
different operationalization of political ideology.

\hypertarget{discussion-1}{%
\subsection{Discussion}\label{discussion-1}}

\hypertarget{general-discussion}{%
\section{General Discussion}\label{general-discussion}}

\begin{itemize}
\tightlist
\item
  Summary: No evidence for identity-based double standards, but some
  evidence for ideology-based double standards
\item
  Two plausible interpretations: (1) conservatives oppose controversial
  protest (no matter the end) \emph{or} (2) the end justifies the mean
\item
  Re (1) it's plausible but what about recent events? (see public
  polling from grant)
\item
  Re (2) why is there no asymmetry?
\item
  Outline the-end-justifies-the-mean idea, highlighting that just
  supporting a cause is not enough, it needs moralization
\item
  Discuss why high-status group are more accepting of controversial
  protest actions
\item
  Discuss limitations (two cultures, two issues, liberal samples;
  system-defending vs.~system-supporting actions; what about other
  conservative causes?)
\item
  Invite researchers to use our paradigm
\end{itemize}

\refsection

\begingroup

\noindent \setlength{\parindent}{-0.5in} \setlength{\leftskip}{0.5in}
\small

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-abrams_double_2013}{}}%
Abrams, D., Randsley de Moura, G., \& Travaglino, G. A. (2013). A double
standard when group members behave badly: {Transgression} credit to
ingroup leaders. \emph{Journal of Personality and Social Psychology},
\emph{105}(5), 799--815. \url{https://doi.org/10.1037/a0033600}

\leavevmode\vadjust pre{\hypertarget{ref-becker_virtual_2012}{}}%
Becker, J. C. (2012). Virtual special issue on theory and research on
collective action in the {European} {Journal} of {Social} {Psychology}
{[}{Editorial}{]}. \emph{European Journal of Social Psychology},
\emph{42}(1), 19--23. \url{https://doi.org/10.1002/ejsp.1839}

\leavevmode\vadjust pre{\hypertarget{ref-becker_committed_2011}{}}%
Becker, J. C., Tausch, N., Spears, R., \& Christ, O. (2011). Committed
dis(s)idents: {Participation} in radical collective action fosters
disidentification with the broader in-group but enhances political
identification. \emph{Personality and Social Psychology Bulletin},
\emph{37}(8), 1104--1116. \url{https://doi.org/10.1177/0146167211407076}

\leavevmode\vadjust pre{\hypertarget{ref-brewer_social_2007}{}}%
Brewer, M. B. (2007). The social psychology of intergroup relations:
Social categorization, ingroup bias, and outgroup prejudice. In A. W.
Kruglanski \& E. T. Higgins (Eds.), \emph{Social psychology: Handbook of
basic principles} (2nd ed., pp. 695--715). Guilford Press.

\leavevmode\vadjust pre{\hypertarget{ref-burkner_bayesian_2021}{}}%
Bürkner, P.-C. (2021). Bayesian item response modeling in {R} with brms
and {Stan}. \emph{Journal of Statistical Software}, \emph{100}(5).
\url{https://doi.org/10.18637/jss.v100.i05}

\leavevmode\vadjust pre{\hypertarget{ref-burkner_modelling_2020}{}}%
Bürkner, P.-C., \& Charpentier, E. (2020). Modelling monotonic effects
of ordinal predictors in {Bayesian} regression models. \emph{British
Journal of Mathematical and Statistical Psychology}, \emph{73}(3),
420--451. \url{https://doi.org/10.1111/bmsp.12195}

\leavevmode\vadjust pre{\hypertarget{ref-demars_irt_2010}{}}%
DeMars, C. (2010). \emph{Item response theory}. Oxford University Press.

\leavevmode\vadjust pre{\hypertarget{ref-dunivin_black_2022}{}}%
Dunivin, Z. O., Yan, H. Y., Ince, J., \& Rojas, F. (2022). Black {Lives}
{Matter} protests shift public discourse. \emph{Proceedings of the
National Academy of Sciences}, \emph{119}(10), e2117320119.
\url{https://doi.org/10.1073/pnas.2117320119}

\leavevmode\vadjust pre{\hypertarget{ref-endevelt_everyone_2021}{}}%
Endevelt, K., Schori-Eyal, N., \& Halperin, E. (2021). Everyone should
get the same, but we should get more: Group entitlement and intergroup
moral double standard. \emph{Group Processes \& Intergroup Relations},
\emph{24}(3), 350--370. \url{https://doi.org/10.1177/1368430219896618}

\leavevmode\vadjust pre{\hypertarget{ref-feinberg_activists_2020}{}}%
Feinberg, M., Willer, R., \& Kovacheff, C. (2020). The activist's
dilemma: {Extreme} protest actions reduce popular support for social
movements. \emph{Journal of Personality and Social Psychology},
\emph{119}(5), 1086--1111. \url{https://doi.org/10.1037/pspi0000230}

\leavevmode\vadjust pre{\hypertarget{ref-foschi_double_2000}{}}%
Foschi, M. (2000). Double standards for competence: Theory and research.
\emph{Annual Review of Sociology}, \emph{26}(1), 21--42.
\url{https://doi.org/10.1146/annurev.soc.26.1.21}

\leavevmode\vadjust pre{\hypertarget{ref-cmdstanr_2021}{}}%
Gabry, J., \& Cesnovar, R. (2021). \emph{{cmdstanr}: {R} interface to
{CmdStan}} (Version 0.4.0). \url{https://mc-stan.org/cmdstanr}

\leavevmode\vadjust pre{\hypertarget{ref-gelman_why_2012}{}}%
Gelman, A., Hill, J., \& Yajima, M. (2012). Why we (usually) don't have
to worry about multiple comparisons. \emph{Journal of Research on
Educational Effectiveness}, \emph{5}(2), 189--211.
\url{https://doi.org/10.1080/19345747.2011.618213}

\leavevmode\vadjust pre{\hypertarget{ref-gelman_prior_2017}{}}%
Gelman, A., Simpson, D., \& Betancourt, M. (2017). The prior can often
only be understood in the context of the likelihood. \emph{Entropy},
\emph{19, 10}, 555. \url{https://doi.org/10.3390/e19100555}

\leavevmode\vadjust pre{\hypertarget{ref-hewstone_ultimate_1990}{}}%
Hewstone, M. (1990). The {``ultimate attribution error''}? {A} review of
the literature on intergroup causal attribution. \emph{European Journal
of Social Psychology}, \emph{20}(4), 311--335.
\url{https://doi.org/10.1002/ejsp.2420200404}

\leavevmode\vadjust pre{\hypertarget{ref-ho_nature_2015}{}}%
Ho, A. K., Sidanius, J., Kteily, N., Sheehy-Skeffington, J., Pratto, F.,
Henkel, K. E., Foels, R., \& Stewart, A. L. (2015). The nature of social
dominance orientation: {Theorizing} and measuring preferences for
intergroup inequality using the new {SDO}-7 scale. \emph{Journal of
Personality and Social Psychology}, \emph{109}(6), 1003--1028.
\url{https://doi.org/10.1037/pspi0000033}

\leavevmode\vadjust pre{\hypertarget{ref-jost_theory_2020}{}}%
Jost, J. T. (2020). \emph{A theory of system justification}. Harvard
University Press.

\leavevmode\vadjust pre{\hypertarget{ref-jost_role_1994}{}}%
Jost, J. T., \& Banaji, M. R. (1994). The role of stereotyping in
system-justification and the production of false consciousness.
\emph{British Journal of Social Psychology}, \emph{33}(1), 1--27.
\url{https://doi.org/10.1111/j.2044-8309.1994.tb01008.x}

\leavevmode\vadjust pre{\hypertarget{ref-jost_missing_2017}{}}%
Jost, J. T., Becker, J., Osborne, D., \& Badaan, V. (2017). Missing in
(collective) action: Ideology, system justification, and the
motivational antecedents of two types of protest behavior. \emph{Current
Directions in Psychological Science}, \emph{26}(2), 99--108.
\url{https://doi.org/10.1177/0963721417690633}

\leavevmode\vadjust pre{\hypertarget{ref-jost_political_2003}{}}%
Jost, J. T., Glaser, J., Kruglanski, A. W., \& Sulloway, F. J. (2003).
Political conservatism as motivated social cognition.
\emph{Psychological Bulletin}, \emph{129}(3), 339--375.
\url{https://doi.org/10.1037/0033-2909.129.3.339}

\leavevmode\vadjust pre{\hypertarget{ref-kay_complementary_2003}{}}%
Kay, A. C., \& Jost, J. T. (2003). Complementary justice: Effects of
"poor but happy" and "poor but honest" stereotype exemplars on system
justification and implicit activation of the justice motive.
\emph{Journal of Personality and Social Psychology}, \emph{85}(5),
823--837. \url{https://doi.org/10.1037/0022-3514.85.5.823}

\leavevmode\vadjust pre{\hypertarget{ref-lacina_nfl_2020}{}}%
Lacina, B. (2020). \emph{National anthem protests and whites' views of
black NFL players} {[}Unpublished manuscript{]}. Department of Political
Science, University of Rochester.
\url{http://www.bethanylacina.com/docs/Lacina_NFL_paper.pdf}

\leavevmode\vadjust pre{\hypertarget{ref-mendoza_for_2014}{}}%
Mendoza, S. A., Lane, S. P., \& Amodio, D. M. (2014). For members only:
Ingroup punishment of fairness norm violations in the ultimatum game.
\emph{Social Psychological and Personality Science}, \emph{5}(6),
662--670. \url{https://doi.org/10.1177/1948550614527115}

\leavevmode\vadjust pre{\hypertarget{ref-osborne_protesting_2019}{}}%
Osborne, D., Jost, J. T., Becker, J. C., Badaan, V., \& Sibley, C. G.
(2019). Protesting to challenge or defend the system? {A} system
justification perspective on collective action. \emph{European Journal
of Social Psychology}, \emph{49}(2), 244--269.
\url{https://doi.org/10.1002/ejsp.2522}

\leavevmode\vadjust pre{\hypertarget{ref-pinto_membership_2010}{}}%
Pinto, I. R., Marques, J. M., Levine, J. M., \& Abrams, D. (2010).
Membership status and subjective group dynamics: Who triggers the black
sheep effect? \emph{Journal of Personality and Social Psychology},
\emph{99}(1), 107--119. \url{https://doi.org/10.1037/a0018187}

\leavevmode\vadjust pre{\hypertarget{ref-reimer_identity_2022}{}}%
Reimer, N. K., Schmid, K., Hewstone, M., \& Al Ramiah, A. (2022).
Self-categorization and social identification: Making sense of us and
them. In D. Chadee (Ed.), \emph{Theories in social psychology} (2nd ed.,
pp. 273--295). Wiley-Blackwell.

\leavevmode\vadjust pre{\hypertarget{ref-samejima_graded_1997}{}}%
Samejima, F. (1997). Graded response model. In W. J. van der Linden \&
R. K. Hambleton (Eds.), \emph{Handbook of modern item response theory}
(pp. 85--100). Springer.

\leavevmode\vadjust pre{\hypertarget{ref-sharp_nonviolent_1973}{}}%
Sharp, G. (1973). \emph{The politics of nonviolent action}. Porter
Sargent.

\leavevmode\vadjust pre{\hypertarget{ref-shuman_disrupting_2021}{}}%
Shuman, E., Saguy, T., van Zomeren, M., \& Halperin, E. (2021).
Disrupting the system constructively: {Testing} the effectiveness of
nonnormative nonviolent collective action. \emph{Journal of Personality
and Social Psychology}, \emph{121}(4), 819--841.
\url{https://doi.org/10.1037/pspi0000333}

\leavevmode\vadjust pre{\hypertarget{ref-cmdstan_2021}{}}%
Stan Development Team. (2021). \emph{Stan modeling language users guide
and reference manual} (Version 2.27.0). \url{http://mc-stan.org/}

\leavevmode\vadjust pre{\hypertarget{ref-tajfel_integrative_1979}{}}%
Tajfel, H., \& Turner, J. C. (1979). An integrative theory of intergroup
conflict. In W. G. Austin \& S. Worchel (Eds.), \emph{The psychology of
intergroup relations} (pp. 33--48). Brooks/Cole.

\leavevmode\vadjust pre{\hypertarget{ref-teixeira_is_2020}{}}%
Teixeira, C. P., Spears, R., \& Yzerbyt, V. Y. (2020). Is {Martin}
{Luther} {King} or {Malcolm} {X} the more acceptable face of protest?
{High}-status groups' reactions to low-status groups' collective action.
\emph{Journal of Personality and Social Psychology}, \emph{118}(5),
919--944. \url{https://doi.org/10.1037/pspi0000195}

\leavevmode\vadjust pre{\hypertarget{ref-valdesolo_moral_2007}{}}%
Valdesolo, P., \& DeSteno, D. (2007). Moral hypocrisy: Social groups and
the flexibility of virtue. \emph{Psychological Science}, \emph{18}(8),
689--690. \url{https://doi.org/10.1111/j.1467-9280.2007.01961.x}

\leavevmode\vadjust pre{\hypertarget{ref-van_zomeren_building_2016}{}}%
van Zomeren, M. (2016). Building a tower of babel? {Integrating} core
motivations and features of social structure into the political
psychology of political action. \emph{Political Psychology}, \emph{37},
87--114. \url{https://doi.org/10.1111/pops.12322}

\leavevmode\vadjust pre{\hypertarget{ref-wright_responding_1990}{}}%
Wright, S. C., Taylor, D. M., \& Moghaddam, F. M. (1990). Responding to
membership in a disadvantaged group: From acceptance to collective
protest. \emph{Journal of Personality and Social Psychology},
\emph{58}(6), 994--1003.
\url{https://doi.org/10.1037/0022-3514.58.6.994}

\leavevmode\vadjust pre{\hypertarget{ref-yougov_nfl_2017}{}}%
YouGov. (2017). \emph{{HuffPost: NFL} {(Sept. 25--26, 2017)}} {[}Data
set{]}. \url{https://perma.cc/3Q3V-4GC4}

\end{CSLReferences}

\endgroup

\newpage

\setcounter{table}{0}
\renewcommand{\thetable}{A\arabic{table}}

\begin{table*}

\section{Appendix A}

\caption{List of protest actions used in Experiment 1}
\small
\begin{tabularx}{\linewidth}{rXr}
\addlinespace
\toprule
\# & Action & Pr\\
\midrule
1 & participate in a public meeting of representatives and elected officials & 97\%\\
2 & hold meetings to inform the public & 96\%\\
3 & make a public speech & 96\%\\
4 & hold meetings to influence the public & 93\%\\
5 & attend or organise a protest march & 93\%\\
6 & attend or organise a protest rally & 92\%\\
7 & use social media (e.g., Facebook, Twitter, Instagram) to influence the public & 92\%\\
8 & do not buy goods or services from companies who support the bill (consumers' boycott) & 92\%\\
9 & paste up posters with political messages in places where it is allowed and encouraged & 89\%\\
10 & join or form a group of activists who oppose the bill & 87\%\\
11 & refuse to accept honours or awards in protest & 82\%\\
12 & donate to political parties who oppose the bill & 80\%\\
13 & refuse to work (strike) & 80\%\\
14 & pay for adverts on social media (e.g., Facebook, Twitter, Instagram) to influence public opinion & 78\%\\
15 & donate to activist groups who oppose the bill & 77\%\\
16 & visit people in their homes to convince them about the issue (canvassing, door knocking) & 54\%\\
17 & stand or sit in a building and refuse to leave (stand-in, sit-in) & 51\%\\
18 & refuse to honour national symbols and traditions (e.g., refusing to sing the national anthem) until the bill is abandoned & 47\%\\
19 & enter and refuse to leave a building (occupation) & 38\%\\
20 & paste up posters with political messages in places where it is not allowed or encouraged & 35\%\\
21 & disrupt traffic (e.g., blocking roads) & 22\%\\
22 & refuse to cooperate with the police and other government agencies & 19\%\\
23 & mock or insult individuals who support the bill & 14\%\\
24 & spray paint political messages in public places & 14\%\\
25 & deface flags or other national symbols & 12\%\\
\bottomrule
\end{tabularx}
\caption*{\textit{Note.} Actions are ordered by the proportion of participants, across all conditions, who considered it to be an acceptable means of protest in Experiment 1.}

\end{table*}

\setcounter{table}{0}
\renewcommand{\thetable}{B\arabic{table}}

\begin{table*}

\section{Appendix B}

\caption{List of protest actions used in Experiment 2}
\small
\begin{tabularx}{\linewidth}{rXr}
\addlinespace
\toprule
\# & Action & Pr\\
\midrule
1 & make a public speech & 94\%\\
2 & \textit{hand out flyers, leaflets, or pamphlets} & 93\%\\
3 & use social media (e.g., Facebook, Twitter, Instagram) to influence the public & 91\%\\
4 & hold meetings to influence the public & 91\%\\
5 & paste up posters with political messages in places where it is allowed and encouraged & 91\%\\
6 & attend or organise a protest march & 91\%\\
7 & donate to activist groups & 89\%\\
8 & join or form a group of activists & 89\%\\
9 & \textit{wear or display political symbols (e.g., patches, flags, bumper stickers)} & 86\%\\
10 & refuse to buy goods or services from companies that advocate [against/for] defunding the police (boycott) & 85\%\\
11 & pay for adverts on social media (e.g., Facebook, Twitter, Instagram) to influence public opinion & 83\%\\
12 & donate to politicians who advocate [for/against] defunding the police & 82\%\\
13 & refuse to accept honors or awards in protest & 80\%\\
14 & refuse to work (strike) & 61\%\\
15 & visit people in their homes to convince them about the issue (canvassing, door knocking) & 61\%\\
16 & stand or sit in a building and refuse to leave (stand-in, sit-in) & 55\%\\
17 & \textit{attend a protest march even though it might turn violent} & 44\%\\
18 & \textit{attend a protest march even though it might be unlawful} & 42\%\\
19 & \textit{attend a protest march while carrying a firearm (where legal)} & 35\%\\
20 & enter and refuse to leave a building (occupation) & 34\%\\
21 & paste up posters with political messages in places where it is not allowed or encouraged & 29\%\\
22 & \textit{refuse to pay fees, fines, or taxes in protest} & 29\%\\
23 & disrupt traffic (e.g., blocking roads) & 25\%\\
24 & spray paint political messages in public places & 23\%\\
25 & mock or insult individuals who are [against/for] defunding the police & 15\%\\
\bottomrule
\end{tabularx}
\caption*{\textit{Note.} Actions are ordered by the proportion of participants, across all conditions, who considered it to be an acceptable means of protest in Experiment 2. Actions in \textit{italics} replaced actions used in Experiment 1 that were either redundant or did not fit the context of Experiment 2.}

\end{table*}

\end{document}