### 2024-04-16 Documents/supplemental-materials.qmd

---
title: Supplemental Online Materials
subtitle: Double Standards in Judging Collective Action
bibliography: references.bib
csl: apa.csl
execute:
  echo: false
format:
  pdf:
    keep-tex: true
    pdf-engine: xelatex
    template: som-template.tex
---

```{r setup, include = FALSE}

  # Load packages
  library(tidyverse)
  library(tidybayes)
  library(numform)
  library(glue)

  # knitr options
  knitr::opts_chunk$set(echo = FALSE)

```

# Scale Development

## Study 1

In Study 1, we compiled a list of collective actions from participants' responses and other sources. To that end, we recruited 60 participants from the Prolific subject pool, all of whom were citizens of the UK or the US. To increase the socioeconomic diversity of our sample, we recruited 30 non-students without a university degree, 15 non-students with a university degree, and 15 current university students. Participants first read an accessible description of collective action:

> Society is not only made up of individuals, but consists of 'social groups' to which these individuals belong. Each person lives in a place, has a job (or not) at an organisation, is a fan of a specific sports club,  has a religion, or belongs to any number of other such groups. Individuals often act in ways to promote the interests of the social groups to which they belong.

> For example, the workers in a factory might want to get paid more. They might go on a strike to reach that goal. Students at a university might want to prevent an increase in tuition fees. They might hand out flyers or occupy a building to reach that goal. Other groups might want to protest against police violence. They might block traffic on a road to make others aware of this issue.

\noindent Participants were then asked to name at least five (and up to ten) actions that fit that description and were encouraged to think of actions that vary in how acceptable or unacceptable they are (in their opinion):

> Please think of other actions that members of a social group (or social groups as a whole) might take to promote their group's interests or goals. Please name at least 5 actions that fit the description above. You can name up to 10 actions if you want to. Please try to think of actions that vary in how acceptable or unacceptable they are (in your opinion).

\noindent After naming at least five actions, participants sorted their responses into actions that, across a range of situations, were "always acceptable", "sometimes acceptable", and "never acceptable":

> On the left, you see the actions you have named on the previous page. Think about each action. Can you think of situations in which this action is an acceptable means to advance a group's goals or interests? Can you think of situations in which this action is not an acceptable means to advance a group's interests?

\noindent Participants named between 5 and 10 actions ($\textit{Mdn} = 7$). Participants named more actions that they considered always acceptable ($\textit{Mdn} = 3$) than actions they considered sometimes ($\textit{Mdn} = 2$) or never ($\textit{Mdn} = 1$) acceptable. We recoded participants' responses into a smaller set of unique collective actions. We then supplemented the resulting list of actions with collective actions from the psychological and political science literature [e.g., @sharp_nonviolent_1973]. This process resulted in 72 actions that we expected to vary in how acceptable most people would find them to be.

## Study 2

In Study 2, we measured how acceptable participants judged the actions from Study 1 to be and applied item response theory to develop an instrument to capture double standards in judging collective action. We recruited 158 participants ($\textit{Mdn} = 30$ years, age range: 18--68 years; 103 women, 52 men, 2 other, 1 prefer not to say) from the Prolific subject pool, all of whom were citizens of the UK or the US. To increase the socioeconomic diversity of our sample, we recruited 80 non-students without a university degree, 37 non-students with a university degree, and 41 current university students. We excluded 15 participants who failed an attention check, leaving a final sample of 143 participants for our analyses.

Participants first read an accessible description of collective action:

> Society is not only made up of individuals, but consists of 'social groups' to which these individuals belong. Each person lives in a place, has a job (or not) at a firm or organisation, has a religion (or not), is part of a political party, or belongs to any number of other groups like these.

> Individuals often act in ways to promote their group's interest. For example, individuals might want people like them to get paid more. Members of a group might take various kinds of actions (for example, go on a strike).

\noindent Participants were then asked to "suppose that one or more members of a group took the following action to advance a cause in their group's interest" and were presented with one of the actions from Study 1. Participants answered several questions about the protest action. First, they rated how disruptive, violent, and extreme they considered this action to be (1 = *not at all*, 4 = *very*). Second, they were asked to think of a range of different causes and circumstances, and to rate how often this action would be an acceptable means for a group to advance one of these causes (1 = *never*, 2 = *rarely*, 3 = *sometimes*, 4 = *often*, 5 = *always*). Finally, they rated how positive or negative they felt, in general, about this action (1 = *very positive*, 5 = *very negative*). Each participant answered these questions for 20 of the 72 actions from Study 1 so that each action was rated by 29--53 participants.

We estimated a graded response model [@burkner_bayesian_2021; @samejima_graded_1997], an item response theory model for ordinal response variables, for participants' ratings of how often an action would be an acceptable form of collective action. For each participant $j$, the model estimated their unique propensity ($\theta_j$) to consider collective actions acceptable in more situations. For each item $i$, the model estimated four acceptability thresholds ($\beta_{ik}$), separating the five response options, and one discrimination parameter ($\alpha_i$) indicating how well the item differentiates between participants with different propensities to consider collective actions acceptable. We estimated the model in *CmdStanR* using similar prior distribution as in Experiment 1. We used an induced Dirichlet prior [@betancourt_ordinal_2019] for the item-specific difficulty thresholds.

Table S1 shows each protest action's information and discrimination parameters as well as the proportions of participants choosing each answer option. Table S2 shows participants' averaged ratings of how acceptable, disruptive, violent, extreme, and negative they considered each action to be. Table S3 shows that all five ratings were strongly correlated across actions and participants.

Figure \ref{fig:s1} shows the scale information curves for scales composed of, respectively, the 1--72 most informative and discriminating protest actions. We selected the 25 most informative actions to construct the final scale to be used in Experiments 1--3. While adding more actions to the scale would have improved the information provided by the instrument, the added benefit of each additional action included in the scale would have been small (see Figure \ref{fig:s1}). We limited the number of protest actions to 25 so as to not overwhelm participants and to prevent satisficing. Tables A1, B1, and C1 list the protest actions used in Experiments 1--3.

```{=latex}
\begin{figure*}[!t]
\centering
\caption{Scale information curves for scales composed of the 1--72 most informative and discriminating protest actions}
\includegraphics[scale=1]{../Scale Development/figures/figure-s1}
\caption*{\textit{Note.} The information curve for the final scale of 25 actions is highlighted in red. As in Figure 1, $\theta_{(z)}$ follows a standard normal distribution and is shown as both $z$-scores (bottom) and percentiles (top).}
\label{fig:s1}
\end{figure*}
```

# Experiment 1

## Method

### Procedure

We used a two-step process to recruit participants who satisfied our preregistered inclusion criteria.

First, we recruited a larger pool of 875 participants. Of these, 475 participants did not have a university degree and placed themselves on the bottom three ranks of the subjective socio-economic status ladder (*lower-status group*) and 400 participants had at least an undergraduate degree and placed themselves on the top four ranks of the subjective socio-economic status ladder (*higher-status group*). Participants completed a screening survey in which they, among other questions, answered whether they considered either their current job or the jobs they had had in the past or would have in the future to be a working-class job or a middle-class/professional job (1 = *working-class job*, 2 = *middle-class/professional job*, 3 = *neither*). Before answering this question, participants read the following explanation:

> The workforce is often divided into two kinds of jobs. What we call working-class jobs are jobs done by skilled, semi-skilled, unskilled manual workers or by casual workers. These jobs do not usually require a university degree. What we call middle-class or professional jobs are administrative, managerial, or other jobs that usually require a university degree.

\noindent As preregistered, we excluded participants from the lower-status group who had responded "middle-class/professional job" or "neither" to this question; we excluded participants from the higher-status group who had responded "working-class job" or "neither" to this question. This left 687 participants for the next step of the selection procedure.

Second, we set out to recruit 500 participants from the remaining 687 participants, 250 from the lower-status group and 250 from the higher-status group. After giving their informed consent, participants were randomly assigned to read a vignette about a government bill affecting either people in working-class jobs (*lower-status protesters*) or people in professional jobs (*higher-status protesters*). Participants in both conditions were instructed to carefully read the vignette and to try to imagine what it would be like if this situation was real. Participants in the lower-status protesters condition then read the following introduction:

> The government, though not necessarily the current government, is going to introduce a bill that will mostly affect people in working-class jobs. Working-class jobs, in this case, are jobs done by skilled, semi-skilled, unskilled manual workers or by casual workers. These are jobs that do not usually require a university degree. Other jobs are unlikely to be affected.

\noindent Participants in the higher-status protesters condition instead read the following introduction:

> The government, though not necessarily the current government, is going to introduce a bill that will mostly affect people in professional jobs. Professional jobs are administrative, managerial, or other jobs that usually require a university degree. Other jobs are unlikely to be affected.

\noindent Participants in the two conditions then read this almost identically worded paragraph:

>This government measure would make it easier for companies to hire workers during economic growth and to lay off workers during an economic crisis. As a consequence, companies would be able to fire employees with little notice and without giving a reason. Trade unions are opposed to the measure. They argue that the bill would compromise job security, and prevent employees from challenging harassment or other abuse without the fear of being fired. People in [working-class/professional] jobs are particularly at risk, and there is a rise in tension and outrage among them.

After reading the vignette, participants responded to the other measures (in the order in which they are presented in the main text). On the final page, we reminded participants that they had read about a bill that the government planned to introduce and that some people objected to. As an attention check, we then asked participants to recall who the people most affected by this measure were. As preregistered, we excluded all participants who did not respond with an accurate description of the group they had read about.

# Experiment 3

## Method

### Measures

We included additional measures not used in the analyses reported in this paper.

We included a three-item measure of moral conviction [@ryan_reconsidering_2014] that assessed to what extent participants' feelings about restricting or banning abortion were, for example, connected to their core moral beliefs or convictions (1 = *not at all*, 5 = *very much*).

We measured gender-related system-justifying beliefs with eight items [adapted from @jost_exposure_2005], for example, "in general, relations between men and women are fair" (1 = *strongly disagree*, 7 = *strongly agree*; McDonald's $\omega = .93$). A confirmatory factor analysis model in which all items loaded onto a single factor showed acceptable fit, $\chi^2 (20) = 111.90; \text{CFI} = 0.97; \text{TLI} = 0.95; \text{RMSEA} = 0.09, [0.08, 0.11]$.

We measured moral concerns with the 36-item moral foundations questionnaire [MFQ-2, @atari_morality_2022] which assesses to what extent participants endorse concerns about care, equality, proportionality, loyalty, authority, and purity (e.g., "I believe chastity is an important virtue."; 1 = *does not describe me at all*, 5 = *describes me extremely well*). We embedded three further attention checks within the questionnaire (e.g., "To show that you are paying attention and giving your best effort, please select 'moderately describes me'.").

<!-- # Reanalysis of Teixeira et al.'s [-@teixeira_white_2022] experiment -->

<!-- At the height of the 2020 Black Lives Matter protests, @teixeira_white_2022 exposed White Americans ($N = 399$) to photos of either peaceful or destructive Black Lives Matter protests and asked them to rate how legitimate (i.e., fair, reasonable, legitimate, and justified) they thought the pictured protests were. As such, their findings make for an interesting comparison to our findings. @teixeira_white_2022, however, reported results from a linear model estimating the three-way interaction between the experimental manipulation, belief in systemic racial injustice, and group identification while controlling for political orientation. This analysis did not provide a direct comparison to our findings.  -->

<!-- To compare theirs to our findings, we instead estimated a series of simpler linear regression models, using the brms R package [@burkner_brms_2017; @burkner_advanced_2018] and weakly informative Student-*t* $(3, 0, 2.5)$ prior distributions, with participants' ratings of how legitimate the pictured protests were as the outcome variable.^[We thank the authors for making their data available online ([https://osf.io/sud2y/](https://osf.io/sud2y/)).] First, we found that, as expected, participants considered peaceful protests far more legitimate than destructive protests ($\text{Cohen's}~d = 1.14, [0.98, 1.30]$). Second, we found that more conservative participants rated destructive protests ($\beta_{xy} = -0.59, [-0.68, -0.49]$) and, to a lesser extent, peaceful protests ($\beta_{xy} = -0.20, [-0.30, -0.10]$) to be less legitimate than more liberal participants. Third, we found that participants who identified more strongly with their (White American) ethnic identity rated destructive protests ($\beta_{xy} = -0.35, [-0.45, -0.24]$) and, to a lesser extent, peaceful protests ($\beta_{xy} = -0.10, [-0.21, 0.01]$) to be less legitimate than participants who identified less strongly with their (White American) ethnic identity. Finally, we found that participants' group identification was no longer associated with how legitimate they rated peaceful ($\beta_{xy} = -0.02, [-0.13, 0.05]$) and destructive ($\beta_{xy} = -0.06, [-0.17, 0.05]$) protests to be after controlling for their political orientation. We refer to findings from our reanalysis of Teixeira et al.'s [-@teixeira_white_2022] experiment in the main text. -->

# Reanalyses Without Exclusions

We replicated the analyses for Experiments 1-3 using the complete data without excluding participants based on the (preregistered) exclusion criteria.

## Experiment 1

```{=latex}
\begin{figure*}[!t]
\centering
\caption{Results from the preregistered (\textbf{A}) and non-preregistered (\textbf{B}) analyses for Experiment 1 without exclusions}
\includegraphics[scale=1]{../Experiment 1/figures/figure-s2}
\caption*{\textit{Note.} HS = Higher Status, LS = Lower Status. $\theta_{(z)}$ is the \textit{z}-standardized tendency to consider more controversial actions acceptable means of protest. (\textbf{A}) $p$ and $n$ are, respectively, the predicted proportion and number of actions a participant would consider acceptable means of protest in each condition. Bars enclose the 95\% most plausible estimates. (\textbf{B}) Ribbons enclose, from the darkest to the lightest shade, the 50\%, 80\%, 95\%, and 99\% most plausible estimates.}
\label{fig:s2}
\end{figure*}
```

Figure \ref{fig:s2} (A) shows estimates for each combination of the protesters' and the participants' group membership. Overall, we found that participants' responses depended on both the protesters' and the participants' group status---but not in the directions predicted by our hypotheses. Contradicting our first hypothesis, participants did not consider protest actions performed by their ingroup to be more acceptable, on average, than the same actions performed by the relevant outgroup (Cohen's $d = -0.01, [-0.20, 0.16]$). Instead, we found that participants from higher-status backgrounds considered protest actions performed by both lower-status ($d = 0.40, [0.15, 0.66$]) and higher-status ($d = 0.37, [0.10, 0.63]$) protesters to be, on average, more acceptable than participants from lower-status backgrounds. Contradicting our second hypothesis, participants considered protest actions performed by higher-status protesters to be less acceptable, on average, than the same actions performed by lower-status protesters ($d = -0.20, [-0.37, -0.02]$).

Figure \ref{fig:s2} shows (B) the $z$-standardized propensity to consider more controversial protest actions acceptable as a function of the three ideological orientation variables. We found that participants who reported a more right-wing political orientation ($\beta_{xy} = -0.24, [-0.34, -0.15]$) and who expressed more agreement with system-justifying beliefs ($\beta_{xy} = -0.21, [-0.31, -0.12]$) tended to find fewer collective actions to be acceptable. In contrast, we found that, after controlling for the other two variables, social dominance orientation was not associated with participants' judgments about how acceptable various collective actions are ($\beta_{xy} = -0.06, [-0.16, 0.03]$). Overall, these findings suggest that people who are right-wing and endorse system-justifying beliefs tend to find various collective actions to be less acceptable than people who are left-wing and do not endorse system-justifying beliefs.

## Experiment 2

### Preregistered Analyses

```{=latex}
\begin{figure*}[!t]
\centering
\caption{Results from the preregistered analyses for Experiment 2}
\includegraphics[scale=1]{../Experiment 2/figures/figure-s3}
\caption*{\textit{Note.} Against = Protesters oppose defunding the police. For = Protesters support defunding the police. $\theta_{(z)}$ is the \textit{z}-standardized tendency to consider more controversial actions acceptable means of protest. Ribbons enclose, from the darkest to the lightest shade, the 50\%, 80\%, 95\%, and 99\% most plausible estimates.}
\label{fig:s3}
\end{figure*}
```

```{=latex}
\begin{figure*}[!t]
\centering
\caption{Predictions from the preregistered (system justification) and non-preregistered (political orientation) analyses for Experiment 2 without exclusions}
\includegraphics[scale=1]{../Experiment 2/figures/figure-s4}
\caption*{\textit{Note.} Against = Protesters oppose defunding the police. For = Protesters support defunding the police. $\theta_{(z)}$ is the \textit{z}-standardized tendency to consider more controversial actions acceptable means of protest.}
\label{fig:s4}
\end{figure*}
```

Figure \ref{fig:s3} shows results from three preregistered models we estimated to test our hypotheses.

```{r, include = FALSE}

  # Import results
  m1 <- read_rds("../Experiment 2/results/m1_post_without_exclusions.rds")

  # Define contrasts
  m1 <- m1 %>% 
    mutate(
      h1a_contrast = case_when(
        Participant == Protesters ~ +1/8,
        Participant != Protesters ~ -1/8,
      ),
      h1b_contrast = case_when(
        Participant == "Black" & Cause == "For" ~ +1/8,
        Participant == "White" & Cause == "Against" ~ +1/8,
        TRUE ~ -1/8,
      ),
      h2a_contrast = case_when(
        Cause == "For" ~ -1/8,
        Cause == "Against" ~ +1/8
      ),
      nh1_contrast = case_when(
        Participant == "Black" ~ -1/8,
        Participant == "White" ~ +1/8
      ),
      nh2_contrast = case_when(
        Participant == "White" & Cause == "Against" ~ -1/4,
        Participant == "White" & Cause == "For" ~ +1/4,
        Participant == "Black" ~ 0
      ),
      nh3_contrast = case_when(
        Participant == "Black" & Cause == "Against" ~ -1/4,
        Participant == "Black" & Cause == "For" ~ +1/4,
        Participant == "White" ~ 0
      )
    )

  # Calculate contrasts
  m1 <- m1 %>% 
    group_by(.draw) %>% 
    summarize(
      across(
        ends_with("_contrast"),
        ~sum(z * .)
      )
    ) %>% 
    pivot_longer(
      ends_with("_contrast"),
      names_to = "contrast",
      values_to = "d"
    ) %>% 
    group_by(contrast) %>% 
    median_qi(d)

  # Make text
  m1 <- m1 %>% 
    mutate(
      across(
        c(d, .lower, .upper), 
        f_num, digits = 2, retain.leading.zero = T
      ),
      d_text = glue(
        "$d = {d}, [{.lower}, {.upper}]$"
      )
    )

```

Model 1 estimated varying intercepts for the eight conditions to test whether, in line with Hypotheses 1a, 1b, or 2a, participants' responses depended on their own group membership, the protesters' group membership, or the protesters' cause. Contradicting *Hypothesis 1a*, participants did not consider protest actions performed by their ingroup to be more acceptable, on average, than the same actions performed by the relevant outgroup (`r with(m1, d_text[contrast == "h1a_contrast"])`). Contradicting *Hypothesis 1b*, participants did not consider protest actions for a cause that was nominally aligned with their ingroup's interests to be more acceptable, on average, than protest actions for a cause not aligned with their ingroup's interests (`r with(m1, d_text[contrast == "h1b_contrast"])`). Contradicting *Hypothesis 2a*, participants did not consider protest actions for a system-defending cause to be more acceptable, on average, than protest actions for a system-challenging cause (`r with(m1, d_text[contrast == "h2a_contrast"])`). Instead, we found that, on average, White participants considered all protest actions to be more acceptable than Black participants (`r with(m1, d_text[contrast == "nh1_contrast"])`) and both Black (`r with(m1, d_text[contrast == "nh3_contrast"])`) and White (`r with(m1, d_text[contrast == "nh2_contrast"])`) participants considered the same actions to be more acceptable when protests were for, rather than against, defunding the police. As in Experiment 1, we thus found that participants' responses depended on both the participants' and the protesters' group memberships---but not in the directions predicted by our hypotheses.

```{r, include = FALSE}

  # Import results
  m2 <- read_rds("../Experiment 2/results/m2_post_without_exclusions.rds")

  # Hypothesis 2
  m2_h2 <- m2 %>% 
    distinct(.draw, b_sjb_for, b_sjb_against) %>% 
    pivot_longer(
      c(b_sjb_for, b_sjb_against),
      names_to = "Cause",
      names_prefix = "b_sjb_",
      values_to = "b"
    ) %>% 
    group_by(Cause) %>% 
    mutate(
      b = case_when(
        Cause == "for" ~ -b,
        Cause == "against" ~ b
      )
    ) %>% 
    median_qi(b) %>% 
    mutate(
      across(
        c(b, .lower, .upper), 
        f_num, digits = 2, retain.leading.zero = T
      ),
      b_text = glue(
        "$\\beta_{xy} = (b), [(.lower), (.upper)]$",
        .open = "(", .close = ")"
      )
    )

```

Model 2 extended Model 1 by estimating participants' responses as a function of their *z*-standardized endorsement of system-justifying beliefs. As preregistered, we modeled this relationship with two fixed effects, one estimating the effect of system-justifying beliefs on judgments about system-defending protest actions and one estimating their effect on judgments about system-challenging protest actions, and one varying effect estimating its variance across conditions. Supporting *Hypothesis 2b*, participants who *rejected* system-justifying beliefs were *more* likely to consider system-challenging protest actions (for defunding the police) to be acceptable means of protest (`r with(m2_h2, b_text[Cause == "for"])`). We did not, however, find evidence for the ideological symmetry implied by *Hypothesis 2b* since participants who *endorsed* system-justifying beliefs were *not* more likely to consider system-defending protest actions (against defunding the police) to be acceptable means of protest (`r with(m2_h2, b_text[Cause == "against"])`).

Figure \ref{fig:s4} shows the estimated pattern of condition-wise differences underlying those fixed effects. Participants who endorsed system-justifying beliefs tended to consider system-challenging and system-defending protest actions to be equally unacceptable. In contrast, participants who rejected system-justifying beliefs evinced ideology-based double standards: In line with *Hypothesis 2b*, they considered the same protest actions to be more acceptable when the protesters challenged the system or, to a lesser extent, when the protesters were from the disadvantaged group.

```{r, include = FALSE}

  # Import results
  m3 <- read_rds("../Experiment 2/results/m3_post_without_exclusions.rds")

  # Hypothesis 2
  m3_h3 <- m3 %>% 
    distinct(.draw, b_sup) %>%
    rename(b = b_sup) %>% 
    median_qi(b) %>% 
    mutate(
      across(
        c(b, .lower, .upper), 
        f_num, digits = 2, retain.leading.zero = T
      ),
      b_text = glue(
        "$\\beta_{y} = (b), [(.lower), (.upper)]$",
        .open = "(", .close = ")"
      )
    )

```

Model 3 extended Model 1 by estimating participants' responses as a function of their self-reported support for the cause of the protest. We recoded participants' responses to create a predictor variable that encoded support for defunding the police when protesters supported defunding the police and opposition to defunding the police when protesters opposed defunding the police. As preregistered, we modeled this relationship as a monotonic effect [@burkner_modelling_2020] that estimated the average change in the outcome variable across predictor categories as well as how much of this change occurred between each of the four pairs of adjacent predictor categories.^[Our model assigned a Dirichlet prior, $\alpha = {1, 1, 1, 1}$, to the proportions of the overall change that was expected to occur between each of the four pairs of predictor categories.] Contradicting *Hypothesis 3*, participants who supported the protesters' cause did not, on average, consider the same protest actions to be more acceptable than participants who opposed the protesters' cause (`r with(m3_h3, b_text)`). Ergo, our results did not support the alternative hypothesis that ideology-based double standards can be reduced to support for, or opposition to, the cause of a protest.

### Non-Preregistered Analyses

```{r, include = FALSE}

  # Import results
  m5 <- read_rds("../Experiment 2/results/m5_post_without_exclusions.rds")

  # Define contrasts
  m5 <- m5 %>% 
    mutate(
      h1a_contrast = case_when(
        Participant == Protesters ~ +1/8,
        Participant != Protesters ~ -1/8
      ),
      h1b_contrast = case_when(
        Participant == "Black" & Cause == "For" ~ +1/8,
        Participant == "White" & Cause == "Against" ~ +1/8,
        TRUE ~ -1/8,
      )
    )

  # Calculate contrasts
  m5 <- m5 %>% 
    group_by(.draw, x_ide) %>% 
    summarize(
      across(
        ends_with("_contrast"),
        ~sum(z * .)
      )
    ) %>% 
    pivot_longer(
      ends_with("_contrast"),
      names_to = "contrast",
      values_to = "d"
    ) %>% 
    group_by(x_ide, contrast) %>% 
    median_qi(d)

  # Make text
  m5 <- m5 %>% 
    mutate(
      across(
        c(d, .lower, .upper), 
        f_num, digits = 2, retain.leading.zero = T
      ),
      d_text = glue(
        "$d = {d}, [{.lower}, {.upper}]$"
      )
    )

```

We explored whether group identification moderated how the participants' and the protesters' group memberships affected participants' responses. To that end, we extended Model 1 by estimating participants' responses as a function of their *z*-standardized identification with their racial ingroup. As in Model 2, we modeled this relationship with a fixed and a varying effect. We found, however, that even participants who strongly identified with their racial ingroup (+1*SD*) did not consider protest actions performed by their ingroup to be more acceptable than the same actions performed by the relevant outgroup (`r with(m5, d_text[contrast == "h1a_contrast" & x_ide == 1L])`) and did not consider protest actions for a cause that was aligned with their ingroup's interests to be more acceptable than protest actions for a cause not aligned with their ingroup's interests (`r with(m5, d_text[contrast == "h1b_contrast" & x_ide == 1L])`). Our non-preregistered analyses thus suggested that group identification did not moderate group differences in judgments about collective action by ingroup and outgroup members.

```{r, include = FALSE}

  # Import results
  m4 <- read_rds("../Experiment 2/results/m4_post_without_exclusions.rds")

  # Hypothesis 2
  m4_exp <- m4 %>% 
    distinct(.draw, b_pol_for, b_pol_against) %>% 
    pivot_longer(
      c(b_pol_for, b_pol_against),
      names_to = "Cause",
      names_prefix = "b_pol_",
      values_to = "b"
    ) %>% 
    group_by(Cause) %>% 
    mutate(
      b = case_when(
        Cause == "for" ~ -b,
        Cause == "against" ~ b
      )
    ) %>% 
    median_qi(b) %>% 
    mutate(
      across(
        c(b, .lower, .upper), 
        f_num, digits = 2, retain.leading.zero = T
      ),
      b_text = glue(
        "$\\beta_{xy} = (b), [(.lower), (.upper)]$",
        .open = "(", .close = ")"
      )
    )

```

We also explored whether we would find ideology-based double standards when operationalizing ideology as political orientation instead of as system justification. To that end, we reran Model 2 with political orientation as the *z*-standardized predictor variable. Our results mirrored the preregistered analyses: More liberal participants were more likely to consider protest actions for defunding the police to be acceptable (`r with(m4_exp, b_text[Cause == "for"])`) but more conservative participants were not more likely to consider protest actions against defunding the police to be acceptable (`r with(m4_exp, b_text[Cause == "against"])`). As Figure \ref{fig:s4} shows, conservative participants tended to consider all protest actions to be equally unacceptable while liberal participants considered the same protest actions to be more acceptable when protesters rallied around a progressive cause. Our non-preregistered analyses thus replicated the ideology-based double standards from the preregistered analyses with a different operationalization of ideology.

## Experiment 3

```{=latex}
\begin{figure*}[!t]
\centering
\caption{Results from the preregistered analyses for Experiment 3 without exclusions}
\includegraphics[scale=1]{../Experiment 3/figures/figure-s5}
\caption*{\textit{Note.} Results show the estimated \textit{z}-standardized tendency ($\theta_{(z)}$) to consider more controversial actions acceptable means to protest for or against restricting abortion as a function of the participants' ideology and gender.}
\label{fig:s5}
\end{figure*}
```

### Preregistered Analyses

```{r, include = FALSE}

  # Import results
  m1 <- read_rds("../Experiment 3/results/m1_post_without_exclusions.rds")

  # Define contrasts
  m1 <- m1 %>% 
    mutate(
      h1_contrast = case_when(
        participant_gender == "Woman" & protesters_cause == "Against" ~ +1L,
        TRUE ~ -1L
      ),
      h2a_contrast = case_when(
        participant_ideology == "Conservative" & protesters_cause == "For" ~ +1L,
        participant_ideology == "Liberal" & protesters_cause == "Against" ~ +1L,
        TRUE ~ -1L
      ),
      h2b_contrast = case_when(
        participant_ideology == "Liberal" & protesters_cause == "Against" ~ +1L,
        TRUE ~ -1L
      ),
      nh1_contrast = case_when(
        participant_ideology == "Conservative" & protesters_cause == "For" ~ +1L,
        participant_ideology == "Conservative" & protesters_cause == "Against" ~ -1L,
        TRUE ~ NA_integer_
      ),
      nh2_contrast = case_when(
        participant_ideology == "Liberal" & protesters_cause == "For" ~ -1L,
        participant_ideology == "Liberal" & protesters_cause == "Against" ~ +1L,
        TRUE ~ NA_integer_
      ),
      nh3_contrast = case_when(
        participant_ideology == "Conservative" & participant_gender == "Man" & protesters_cause == "For" ~ +1L,
        participant_ideology == "Conservative" & participant_gender == "Man" & protesters_cause == "Against" ~ -1L,
        TRUE ~ NA_integer_
      ),
      nh4_contrast = case_when(
        participant_ideology == "Conservative" & participant_gender == "Woman" & protesters_cause == "For" ~ +1L,
        participant_ideology == "Conservative" & participant_gender == "Woman" & protesters_cause == "Against" ~ -1L,
        TRUE ~ NA_integer_
      ),
      nh5_contrast = case_when(
        participant_ideology == "Liberal" & participant_gender == "Man" & protesters_cause == "For" ~ -1L,
        participant_ideology == "Liberal" & participant_gender == "Man" & protesters_cause == "Against" ~ +1L,
        TRUE ~ NA_integer_
      ),
      nh6_contrast = case_when(
        participant_ideology == "Liberal" & participant_gender == "Woman" & protesters_cause == "For" ~ -1L,
        participant_ideology == "Liberal" & participant_gender == "Woman" & protesters_cause == "Against" ~ +1L,
        TRUE ~ NA_integer_
      )
    )

  # Calculate contrasts
  m1 <- m1 %>% 
    pivot_longer(
      ends_with("_contrast"),
      names_to = "contrast",
      values_to = "direction",
      values_drop_na = TRUE
    ) %>% 
    group_by(.draw, contrast, direction) %>% 
    summarize(z = mean(z)) %>% 
    summarize(d = sum(z * direction)) %>% 
    pivot_wider(
      names_from = contrast,
      values_from = d
    ) %>% 
    mutate(nh7_contrast = nh2_contrast - nh1_contrast) %>% 
    pivot_longer(
      ends_with("_contrast"),
      names_to = "contrast",
      values_to = "d"
    ) %>% 
    group_by(contrast) %>% 
    median_qi(d)

  # Make text
  m1 <- m1 %>% 
    mutate(
      across(
        c(d, .lower, .upper), 
        f_num, digits = 2, retain.leading.zero = T
      ),
      d_text = glue(
        "$d = {d}, [{.lower}, {.upper}]$"
      )
    )

```

Figure \ref{fig:s5} shows the results of our preregistered analyses without exclusions.

As in Experiments 1 and 2, we found evidence for ideology-based double standards: Conservative participants considered the same protest actions to be more acceptable, on average, when protesters *supported* restricting abortion (`r with(m1, d_text[contrast == "nh1_contrast"])`) while liberal participants considered the same protest actions to be more acceptable when protesters *opposed* restricting abortion (`r with(m1, d_text[contrast == "nh2_contrast"])`). This double standard was more pronounced among liberal participants (`r with(m1, d_text[contrast == "nh7_contrast"])`). In this way, Experiment 3 provided evidence for both ideological symmetry and asymmetry in judging collective action for and against restricting abortion.

We found some evidence for identity-based double standards: Female participants considered protest actions *against* restricting abortion to be more acceptable, on average, than male participants or protest actions *for* restricting abortion (`r with(m1, d_text[contrast == "h1_contrast"])`). This pattern was, however, overshadowed by stronger ideology-based double standards that were consistent across subsamples: Conservative women (`r with(m1, d_text[contrast == "nh4_contrast"])`) and, to a lesser extent, conservative men (`r with(m1, d_text[contrast == "nh3_contrast"])`) considered the same protest actions to be more acceptable when protesters *supported* restricting abortion and both liberal women (`r with(m1, d_text[contrast == "nh6_contrast"])`) and liberal men (`r with(m1, d_text[contrast == "nh5_contrast"])`) considered the same protest actions to be more acceptable when protesters *opposed* restricting abortion. In this way, Experiment 3 provided more consistent evidence for ideology-based than for identity-based double standards in judging collective action.

### Non-Preregistered Analyses

```{r, include = FALSE}

  # Import results
  ide <- read_rds("../Experiment 3/results/ide_post_without_exclusions.rds")

  # Make text
  ide <- ide %>%
    group_by(.draw, participant_gender, protesters_cause) %>% 
    summarize(
      b = mean(b_ide_kk)
    ) %>% 
    group_by(participant_gender, protesters_cause) %>%
    median_qi(b) %>% 
    mutate(
      across(
        c(b, .lower, .upper),
        f_num, digits = 2, retain.leading.zero = T
      ),
      b_text = glue(
        "$\\beta_{xy} = (b), [(.lower), (.upper)]$",
        .open = "(", .close = ")"
      )
    )

```

We explored whether group identification moderated participants' judgments. We found, however, that gender identification neither affected how women judged protesters opposing (`r with(ide, b_text[participant_gender == "Woman" & protesters_cause == "Against"])`) or supporting (`r with(ide, b_text[participant_gender == "Woman" & protesters_cause == "For"])`) abortion restrictions nor how men judged protesters opposing (`r with(ide, b_text[participant_gender == "Man" & protesters_cause == "Against"])`) or supporting (`r with(ide, b_text[participant_gender == "Man" & protesters_cause == "For"])`) abortion restrictions. Our non-preregistered analyses thus suggested that gender identification did not moderate identity-based double standards in judging collective action concerning reproductive rights.

\newpage

# References

\begingroup

\noindent
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs"></div>

\endgroup

\newpage

```{=latex}
\begin{xltabular}{\linewidth}{rXrrrrrrr}

\caption{Results from Study 2}\\
\toprule
   &        &     &          & \multicolumn{5}{c}{Response $[\%]$} \\ \cmidrule{5-9}
\# & Action & $I$ & $\alpha$ & 1 & 2 & 3 & 4 & 5\\
\midrule
\endfirsthead

\toprule
   &        &     &          & \multicolumn{5}{c}{Response $[\%]$} \\ \cmidrule{5-9}
\# & Action & $I$ & $\alpha$ & 1 & 2 & 3 & 4 & 5\\
\midrule
\endhead

\bottomrule
\addlinespace
\caption*{\textit{Note.} $I$ = Information; $\alpha$ = Discrimination; Response: 1 = \textit{never}, 2 = \textit{rarely}, 3 = \textit{sometimes}, 4 = \textit{often}, 5 = \textit{always}}
\endfoot

\bottomrule
\addlinespace
\caption*{\textit{Note.} $I$ = Information; $\alpha$ = Discrimination; Response: 1 = \textit{never}, 2 = \textit{rarely}, 3 = \textit{sometimes}, 4 = \textit{often}, 5 = \textit{always}}
\endlastfoot

1 & disrupt traffic (e.g., blocking roads) & 7.98 & 1.31 & 41 & 36 & 18 & 5 & 0\\
2 & attend or organise a protest rally & 7.35 & 1.11 & 0 & 7 & 45 & 39 & 9\\
3 & refuse to work (strike) & 6.96 & 1.11 & 2 & 5 & 53 & 33 & 7\\
4 & enter and refuse to leave a building (occupation) & 6.95 & 1.12 & 6 & 52 & 26 & 13 & 3\\
5 & deface flags or other national symbols & 6.83 & 1.21 & 52 & 20 & 20 & 9 & 0\\

6 & refuse to honour national symbols and traditions (e.g., refusing to sing the national anthem) & 6.42 & 1.19 & 15 & 18 & 40 & 15 & 12\\
7 & paste up posters with political messages in places where it is not allowed or encouraged & 6.37 & 1.01 & 14 & 32 & 35 & 19 & 0\\
8 & refuse to cooperate with the police and other government agencies & 6.37 & 1.11 & 22 & 38 & 32 & 5 & 3\\
9 & refuse to accept honours or awards in protest & 6.32 & 1.14 & 11 & 14 & 39 & 23 & 14\\
10 & disrupt public events (e.g., a sports game) with a political message & 6.30 & 1.02 & 24 & 45 & 21 & 10 & 0\\

11 & spray paint political messages in public places & 6.28 & 1.06 & 31 & 39 & 28 & 3 & 0\\
12 & stand or sit in a building and refuse to leave (stand-in, sit-in) & 6.28 & 1.04 & 7 & 41 & 34 & 14 & 5\\
13 & pay for adverts on social media (e.g., Facebook, Twitter, Instagram) to influence public opinion & 6.23 & 0.99 & 0 & 16 & 27 & 39 & 18\\
14 & get involved in the media (e.g., newspapers, radio, television) to influence the public & 6.18 & 0.98 & 0 & 11 & 32 & 36 & 20\\
15 & visit people in their homes to convince them about an issue (canvassing, door knocking) & 6.07 & 1.03 & 9 & 30 & 27 & 27 & 6\\

16 & donate to political parties who support the cause & 6.00 & 1.10 & 6 & 6 & 33 & 31 & 25\\
17 & do not buy goods or services from companies who oppose the cause (consumers' boycott) & 5.99 & 1.06 & 2 & 5 & 39 & 29 & 24\\
18 & refuse to interact with or acknowledge individuals who oppose the cause & 5.89 & 0.97 & 21 & 45 & 29 & 5 & 0\\
19 & attend or organise a protest march & 5.88 & 1.06 & 3 & 3 & 34 & 38 & 22\\
20 & paste up posters with political messages in places where it is allowed and encouraged & 5.85 & 1.00 & 0 & 3 & 20 & 43 & 34\\

21 & refuse to pay rent (rent strike) & 5.82 & 0.97 & 23 & 41 & 28 & 8 & 0\\
22 & use social media (e.g., Facebook, Twitter, Instagram) to influence the public & 5.82 & 1.01 & 5 & 5 & 30 & 45 & 15\\
23 & wear or display political symbols & 5.79 & 1.00 & 3 & 9 & 38 & 32 & 18\\
24 & mock or insult individuals who oppose the cause & 5.73 & 1.11 & 53 & 34 & 3 & 11 & 0\\
25 & write letters to politicians, representatives and elected officials & 5.69 & 1.07 & 0 & 0 & 11 & 37 & 53\\

26 & use social media (e.g., Facebook, Twitter, Instagram) to inform the public & 5.67 & 1.01 & 0 & 3 & 12 & 41 & 44\\
27 & join or form a group of activists & 5.66 & 1.02 & 2 & 12 & 21 & 38 & 26\\
28 & hold meetings to inform the public & 5.65 & 1.01 & 0 & 3 & 11 & 38 & 49\\
29 & donate to activist groups who support the cause & 5.64 & 1.06 & 2 & 9 & 41 & 14 & 34\\
30 & disrupt public services (e.g., shutting down government websites) & 5.61 & 1.01 & 42 & 42 & 17 & 0 & 0\\

31 & refuse to pay fees, fines, and taxes & 5.56 & 0.96 & 9 & 50 & 30 & 7 & 4\\
32 & hold meetings to influence the public & 5.56 & 1.00 & 0 & 3 & 16 & 35 & 46\\
33 & hand out flyers, leaflets, or pamphlets & 5.55 & 1.01 & 0 & 0 & 13 & 45 & 42\\
34 & visit people in their homes to inform them about an issue (canvassing, door knocking) & 5.39 & 0.95 & 9 & 20 & 30 & 32 & 9\\
35 & make a public speech & 5.35 & 1.00 & 2 & 2 & 14 & 46 & 36\\

36 & join or form a political party & 5.28 & 0.96 & 2 & 7 & 19 & 45 & 26\\
37 & refuse to interact with or acknowledge politicians who oppose the cause & 5.27 & 0.95 & 8 & 35 & 38 & 10 & 10\\
38 & disrupt private life of politicians (e.g., protesting outside their home) & 5.24 & 0.92 & 39 & 36 & 23 & 2 & 0\\
39 & get involved in the media (e.g., newspapers, radio, television) to inform the public & 5.23 & 0.93 & 0 & 2 & 11 & 43 & 43\\
40 & pay for adverts in the media (e.g., newspapers, radio, television) to influence public opinion & 5.23 & 0.97 & 5 & 10 & 22 & 42 & 20\\

41 & damage private property (e.g., cars or houses) & 5.17 & 1.03 & 70 & 25 & 5 & 0 & 0\\
42 & refuse service (e.g., in a restaurant or shop) to politicians who oppose the cause & 5.16 & 0.89 & 29 & 37 & 26 & 8 & 0\\
43 & sign or start a petition & 5.10 & 0.96 & 0 & 0 & 8 & 39 & 53\\
44 & mock or insult politicians who oppose the cause & 5.07 & 0.99 & 38 & 31 & 16 & 11 & 4\\
45 & join or form a trade/labor union & 5.07 & 0.99 & 2 & 2 & 14 & 37 & 44\\

46 & donate to charities who support the cause & 5.03 & 0.96 & 0 & 0 & 8 & 39 & 53\\
47 & refuse to eat (hunger strike) & 4.95 & 0.97 & 38 & 24 & 29 & 6 & 3\\
48 & damage public property (e.g., goverment buildings) & 4.83 & 0.98 & 72 & 20 & 5 & 2 & 0\\
49 & spread rumours about politicians who oppose the cause & 4.77 & 1.00 & 69 & 21 & 10 & 0 & 0\\
50 & participate in a public meeting of representatives and elected officials & 4.70 & 1.00 & 3 & 0 & 7 & 47 & 43\\

51 & physically harm oneself (e.g., setting oneself on fire) & 4.70 & 1.03 & 80 & 20 & 0 & 0 & 0\\
52 & boycott an election by not voting or spoiling one's ballot & 4.70 & 0.91 & 14 & 38 & 26 & 12 & 10\\
53 & stand in an election & 4.69 & 0.90 & 2 & 8 & 18 & 35 & 38\\
54 & soil politicians who oppose the cause (e.g., throwing eggs at them) & 4.62 & 1.02 & 70 & 21 & 8 & 0 & 2\\
55 & attack politicians with the intention of harming them (e.g., punching them) & 4.62 & 1.01 & 78 & 22 & 0 & 0 & 0\\

56 & refuse service (e.g., in a restaurant or shop) to individuals who oppose the cause & 4.60 & 0.91 & 32 & 54 & 11 & 0 & 3\\
57 & blackmail individuals who oppose the cause & 4.53 & 0.96 & 75 & 22 & 3 & 0 & 0\\
58 & threaten politicians who oppose the cause with physical harm & 4.37 & 1.00 & 86 & 11 & 3 & 0 & 0\\
59 & attack individuals who oppose the cause with the intention of harming them (e.g., punching them) & 4.36 & 0.95 & 82 & 12 & 6 & 0 & 0\\
60 & damage commerical property (e.g., shop windows) & 4.36 & 1.03 & 88 & 12 & 0 & 0 & 0\\

61 & attack politicians with the intention of killing them (e.g., stabbing them) & 4.30 & 1.01 & 91 & 7 & 2 & 0 & 0\\
62 & bribe politicians, representatives, and other elected officials & 4.24 & 0.95 & 85 & 13 & 2 & 0 & 0\\
63 & voting for candidates/parties & 4.23 & 0.85 & 0 & 5 & 7 & 21 & 67\\
64 & donate to political parties to make them change their position on an issue & 4.21 & 0.90 & 35 & 24 & 30 & 3 & 8\\
65 & blackmail politicians who oppose the cause & 4.20 & 0.88 & 72 & 26 & 3 & 0 & 0\\

66 & attack police officers and other government agents with the intention of killing them (e.g., stabbing them) & 4.20 & 1.02 & 93 & 7 & 0 & 0 & 0\\
67 & attack police officers and other government agents with the intention of harming them (e.g., punching them) & 4.17 & 1.00 & 92 & 5 & 3 & 0 & 0\\
68 & attack individuals who oppose the cause with the intention of killing them (e.g., stabbing them) & 3.98 & 1.00 & 96 & 2 & 2 & 0 & 0\\
69 & threaten individuals who oppose the cause with physical harm & 3.96 & 0.99 & 94 & 6 & 0 & 0 & 0\\
70 & attack members of the public (e.g., by setting off a bomb in a public place) & 3.72 & 0.99 & 98 & 0 & 2 & 0 & 0\\

71 & spread misinformation to influence public opinion & 3.71 & 0.82 & 78 & 19 & 3 & 0 & 0\\
72 & threaten to attack members of the public (e.g., by making a bomb threat) & 3.51 & 0.97 & 98 & 0 & 0 & 2 & 0

\end{xltabular}
```

\newpage

```{=latex}
\setcounter{table}{1}

\begin{xltabular}{\linewidth}{rXrrrrr}

\caption{Ratings from Study 2}\\
\toprule
   &        & \multicolumn{5}{c}{Rating $[M]$} \\ \cmidrule{3-7}
\# & Action & Acce. & Disr. & Viol. & Extr. & Nega.\\
\midrule
\endfirsthead

\toprule
   &        & \multicolumn{5}{c}{Rating $[M]$} \\ \cmidrule{3-7}
\# & Action & Acce. & Disr. & Viol. & Extr. & Nega.\\
\midrule
\endhead

\bottomrule
\addlinespace
\caption*{\textit{Note.} Acce. = Acceptable (1 = \textit{never}, 5 = \textit{always}); Disr. = Disruptive, Viol. = Violent, Extr. = Extreme (1 = \textit{not at all}, 4 = \textit{very}); Nega. = Negative (1 = \textit{very positive}, 5 = \textit{very negative})}
\endfoot

\bottomrule
\addlinespace
\caption*{\textit{Note.} Acce. = Acceptable (1 = \textit{never}, 5 = \textit{always}); Disr. = Disruptive, Viol. = Violent, Extr. = Extreme (1 = \textit{not at all}, 4 = \textit{very}); Nega. = Negative (1 = \textit{very positive}, 5 = \textit{very negative})}
\endlastfoot

1 & disrupt traffic (e.g., blocking roads) & 1.86 & 3.82 & 2.05 & 2.98 & 4.16\\
2 & attend or organise a protest rally & 3.50 & 2.50 & 1.36 & 1.48 & 2.64\\
3 & refuse to work (strike) & 3.37 & 3.35 & 1.09 & 2.07 & 2.60\\
4 & enter and refuse to leave a building (occupation) & 2.55 & 3.32 & 1.48 & 2.42 & 3.19\\
5 & deface flags or other national symbols & 1.85 & 2.98 & 2.30 & 2.98 & 4.15\\

6 & refuse to honour national symbols and traditions (e.g., refusing to sing the national anthem) & 2.92 & 1.85 & 1.23 & 1.62 & 3.20\\
7 & paste up posters with political messages in places where it is not allowed or encouraged & 2.59 & 2.84 & 1.22 & 1.92 & 3.38\\
8 & refuse to cooperate with the police and other government agencies & 2.30 & 3.14 & 2.11 & 2.78 & 3.76\\
9 & refuse to accept honours or awards in protest & 3.14 & 1.84 & 1.07 & 1.59 & 2.86\\
10 & disrupt public events (e.g., a sports game) with a political message & 2.17 & 3.40 & 1.60 & 2.31 & 3.79\\

11 & spray paint political messages in public places & 2.03 & 2.72 & 1.53 & 2.44 & 3.97\\
12 & stand or sit in a building and refuse to leave (stand-in, sit-in) & 2.68 & 3.34 & 1.27 & 2.16 & 3.02\\
13 & pay for adverts on social media (e.g., Facebook, Twitter, Instagram) to influence public opinion & 3.59 & 1.59 & 1.02 & 1.30 & 2.75\\
14 & get involved in the media (e.g., newspapers, radio, television) to influence the public & 3.66 & 1.61 & 1.02 & 1.23 & 2.41\\
15 & visit people in their homes to convince them about an issue (canvassing, door knocking) & 2.91 & 2.76 & 1.06 & 1.36 & 3.33\\

16 & donate to political parties who support the cause & 3.64 & 1.14 & 1.00 & 1.17 & 2.47\\
17 & do not buy goods or services from companies who oppose the cause (consumers' boycott) & 3.68 & 2.07 & 1.05 & 1.29 & 2.27\\
18 & refuse to interact with or acknowledge individuals who oppose the cause & 2.18 & 2.42 & 1.24 & 2.37 & 3.79\\
19 & attend or organise a protest march & 3.72 & 2.56 & 1.31 & 1.56 & 2.41\\
20 & paste up posters with political messages in places where it is allowed and encouraged & 4.09 & 1.31 & 1.00 & 1.06 & 2.17\\

21 & refuse to pay rent (rent strike) & 2.21 & 2.72 & 1.10 & 2.41 & 3.67\\
22 & use social media (e.g., Facebook, Twitter, Instagram) to influence the public & 3.60 & 1.40 & 1.00 & 1.15 & 2.42\\
23 & wear or display political symbols & 3.53 & 1.29 & 1.03 & 1.09 & 2.71\\
24 & mock or insult individuals who oppose the cause & 1.71 & 2.79 & 1.95 & 2.58 & 4.18\\
25 & write letters to politicians, representatives and elected officials & 4.42 & 1.39 & 1.05 & 1.03 & 1.53\\

26 & use social media (e.g., Facebook, Twitter, Instagram) to inform the public & 4.25 & 1.31 & 1.00 & 1.06 & 2.03\\
27 & join or form a group of activists & 3.74 & 1.81 & 1.38 & 1.50 & 2.38\\
28 & hold meetings to inform the public & 4.32 & 1.19 & 1.00 & 1.03 & 1.81\\
29 & donate to activist groups who support the cause & 3.68 & 1.39 & 1.11 & 1.25 & 2.39\\
30 & disrupt public services (e.g., shutting down government websites) & 1.75 & 3.75 & 1.61 & 3.06 & 4.28\\

31 & refuse to pay fees, fines, and taxes & 2.48 & 2.61 & 1.04 & 2.09 & 3.46\\
32 & hold meetings to influence the public & 4.24 & 1.14 & 1.00 & 1.03 & 1.86\\
33 & hand out flyers, leaflets, or pamphlets & 4.29 & 1.47 & 1.03 & 1.05 & 2.00\\
34 & visit people in their homes to inform them about an issue (canvassing, door knocking) & 3.11 & 2.45 & 1.07 & 1.50 & 3.16\\
35 & make a public speech & 4.12 & 1.46 & 1.00 & 1.06 & 1.92\\

36 & join or form a political party & 3.86 & 1.40 & 1.00 & 1.24 & 2.10\\
37 & refuse to interact with or acknowledge politicians who oppose the cause & 2.80 & 1.88 & 1.15 & 1.88 & 3.52\\
38 & disrupt private life of politicians (e.g., protesting outside their home) & 1.89 & 3.36 & 2.07 & 2.77 & 4.07\\
39 & get involved in the media (e.g., newspapers, radio, television) to inform the public & 4.27 & 1.36 & 1.02 & 1.14 & 1.82\\
40 & pay for adverts in the media (e.g., newspapers, radio, television) to influence public opinion & 3.62 & 1.60 & 1.10 & 1.35 & 2.50\\

41 & damage private property (e.g., cars or houses) & 1.34 & 3.77 & 3.64 & 3.86 & 4.75\\
42 & refuse service (e.g., in a restaurant or shop) to politicians who oppose the cause & 2.13 & 2.95 & 1.26 & 2.39 & 3.71\\
43 & sign or start a petition & 4.45 & 1.18 & 1.03 & 1.03 & 1.74\\
44 & mock or insult politicians who oppose the cause & 2.13 & 2.31 & 1.58 & 2.02 & 3.84\\
45 & join or form a trade/labor union & 4.19 & 1.40 & 1.00 & 1.07 & 1.70\\

46 & donate to charities who support the cause & 4.44 & 1.03 & 1.00 & 1.03 & 1.47\\
47 & refuse to eat (hunger strike) & 2.12 & 2.26 & 1.38 & 3.32 & 3.74\\
48 & damage public property (e.g., goverment buildings) & 1.38 & 3.67 & 3.55 & 3.65 & 4.68\\
49 & spread rumours about politicians who oppose the cause & 1.41 & 3.14 & 1.72 & 2.76 & 4.31\\
50 & participate in a public meeting of representatives and elected officials & 4.27 & 1.23 & 1.07 & 1.07 & 1.80\\

51 & physically harm oneself (e.g., setting oneself on fire) & 1.20 & 3.68 & 3.76 & 3.95 & 4.83\\
52 & boycott an election by not voting or spoiling one's ballot & 2.64 & 2.24 & 1.19 & 1.90 & 3.45\\
53 & stand in an election & 3.98 & 1.27 & 1.02 & 1.15 & 1.98\\
54 & soil politicians who oppose the cause (e.g., throwing eggs at them) & 1.43 & 3.28 & 3.11 & 3.19 & 4.47\\
55 & attack politicians with the intention of harming them (e.g., punching them) & 1.22 & 3.65 & 3.81 & 3.84 & 4.81\\

56 & refuse service (e.g., in a restaurant or shop) to individuals who oppose the cause & 1.86 & 3.03 & 1.46 & 2.65 & 4.16\\
57 & blackmail individuals who oppose the cause & 1.28 & 3.33 & 2.56 & 3.61 & 4.72\\
58 & threaten politicians who oppose the cause with physical harm & 1.16 & 3.57 & 3.76 & 3.81 & 4.81\\
59 & attack individuals who oppose the cause with the intention of harming them (e.g., punching them) & 1.24 & 3.76 & 3.88 & 3.88 & 4.79\\
60 & damage commerical property (e.g., shop windows) & 1.12 & 3.85 & 3.79 & 3.79 & 4.91\\

61 & attack politicians with the intention of killing them (e.g., stabbing them) & 1.11 & 3.93 & 4.00 & 4.00 & 4.91\\
62 & bribe politicians, representatives, and other elected officials & 1.17 & 3.02 & 1.28 & 3.04 & 4.85\\
63 & voting for candidates/parties & 4.51 & 1.19 & 1.09 & 1.07 & 1.63\\
64 & donate to political parties to make them change their position on an issue & 2.24 & 2.03 & 1.05 & 1.68 & 3.68\\
65 & blackmail politicians who oppose the cause & 1.31 & 3.15 & 2.15 & 3.54 & 4.64\\

66 & attack police officers and other government agents with the intention of killing them (e.g., stabbing them) & 1.07 & 4.00 & 4.00 & 4.00 & 4.98\\
67 & attack police officers and other government agents with the intention of harming them (e.g., punching them) & 1.11 & 3.81 & 3.95 & 3.95 & 4.95\\
68 & attack individuals who oppose the cause with the intention of killing them (e.g., stabbing them) & 1.06 & 3.94 & 4.00 & 4.00 & 4.92\\
69 & threaten individuals who oppose the cause with physical harm & 1.06 & 3.72 & 3.92 & 3.97 & 4.92\\
70 & attack members of the public (e.g., by setting off a bomb in a public place) & 1.05 & 3.90 & 3.92 & 3.92 & 4.92\\

71 & spread misinformation to influence public opinion & 1.24 & 3.41 & 1.41 & 2.95 & 4.76\\
72 & threaten to attack members of the public (e.g., by making a bomb threat) & 1.07 & 3.83 & 3.83 & 3.85 & 4.95

\end{xltabular}
```

\newpage

```{=latex}
\setcounter{table}{2}

\begin{table}[t!]
\caption{Correlations between ratings in Study 2}
\centering
\begin{tabular}{rlrrrrr}
\toprule
\# & Rating & Acceptable & Disruptive & Violent & Extreme & Negative\\
\midrule1 & Acceptable &  & -.69 & -.64 & -.77 & -.89\\
2 & Disruptive & -.69 &  & .66 & .78 & .70\\
3 & Violent & -.64 & .66 &  & .79 & .64\\
4 & Extreme & -.77 & .78 & .79 &  & .77\\
5 & Negative & -.89 & .70 & .64 & .77 & \\
\bottomrule
\addlinespace
\end{tabular}
\caption*{\textit{Note.} Acceptable (1 = \textit{never}, 5 = \textit{always}); Disruptive, Violent, Extreme (1 = \textit{not at all}, 4 = \textit{very}); Negative (1 = \textit{very positive}, 5 = \textit{very negative})}
\end{table}
```

```{=latex}
\setcounter{table}{3}

\begin{table}[t!]
\caption{Correlations between variables in Experiment 1}
\centering
\begin{tabular}{rlrrrr}
\toprule
\# & Variable              &    1 &    2 &    3 &    4 \\ \midrule
 1 & Proportion            &      & -.31 & -.16 & -.23 \\
 2 & Political Orientation & -.31 &      &  .41 &  .41 \\
 3 & Social Dominance      & -.16 &  .41 &      &  .29 \\
 4 & System Justification  & -.23 &  .41 &  .29 &      \\ \bottomrule
\addlinespace
\end{tabular}
\caption*{\textit{Note.} Proportion of actions considered acceptable (0--1); Political Orientation (1 = \textit{left}, 7 = \textit{right}); Social Dominance Orientation (\textit{factor score}); System-Justifying Beliefs (\textit{factor score})}
\end{table}
```

```{=latex}
\setcounter{table}{4}

\begin{table}[t!]
\caption{Correlations between variables in Experiment 2}
\centering
\begin{tabular}{rlrrrrr}
\toprule
\# & Variable              &    1 &    2 &    3 &    4 &    5 \\ \midrule
 1 & Proportion            &      & -.21 &  .29 & -.23 & -.07 \\
 2 & System Justification  & -.21 &      & -.45 &  .47 &  .04 \\
 3 & Policy Support        &  .29 & -.45 &      & -.49 &  .00 \\
 4 & Political Orientation & -.23 &  .47 & -.49 &      &  .03 \\
 5 & Identification        & -.07 &  .04 &  .00 &  .03 &      \\ \bottomrule
\addlinespace
\end{tabular}
\caption*{\textit{Note.} Proportion of actions considered acceptable (0--1); System-Justifying Beliefs (\textit{factor score}); Policy Support (1 = \textit{strongly oppose}, 5 = \textit{strongly support}); Political Orientation (1 = \textit{liberal}, 7 = \textit{conservative}); Racial Identification (1 = \textit{not at all}, 7 = \textit{very much})}
\end{table}
```

```{=latex}
\setcounter{table}{5}

\begin{table}[t!]
\caption{Correlations between variables in Experiment 3}
\centering
\begin{tabular}{rlrrrrr}
\toprule
\# & Variable              &    1 &    2 &    3 &    4 &    5 \\ \midrule
 1 & Proportion            &      & -.27 & -.22 & -.32 & -.11 \\
 2 & System Justification  & -.27 &      &  .61 &  .78 &  .16 \\
 3 & Policy Support        & -.22 &  .61 &      &  .74 &  .16 \\
 4 & Political Orientation & -.32 &  .78 &  .74 &      &  .20 \\
 5 & Identification        & -.11 &  .16 &  .16 &  .20 &      \\ \bottomrule
\addlinespace
\end{tabular}
\caption*{\textit{Note.} Proportion of actions considered acceptable (0--1); System-Justifying Beliefs (\textit{factor score}); Policy Support (1 = \textit{strongly oppose}, 5 = \textit{strongly support}); Political Orientation (1 = \textit{liberal}, 7 = \textit{conservative}); Gender Identification (1 = \textit{not at all}, 7 = \textit{very much})}
\end{table}
```